# Boiler Plate

```{r}
library(readxl)
library(dplyr)
#library(ggplot2)
#library(reshape2)
#library(stringr)
library(tibble)
library(vegan)
```

# Functions

```{r}
`%notin%` <- function (x, y) {!(x %in% y)}
```

# Load data sets

## Process metadata

```{r}
# When the data was gross, process it, set new variables, and isolate the important ones
# metadata <- read_excel(path = "../Metadata_clean.xlsx", 
#                        sheet = 2, 
#                        col_types = c(rep("guess", 7), "date", rep("guess", 12))) %>%
#   filter(Fastq_name %notin% c("79_S70","86_S76", "88_S106")) %>% 
#   as.data.frame()
#          # Drop sample 79_S70 which did not have enough reads for analysis
#          # Also drop sample 86_S76 whose ID is unknown
# metadata$Group[metadata$Group == "Captive"] = "Ex-situ"
# metadata$Specific_Group[metadata$Specific_Group == "Captive"] = "Ex-situ"
# 
# # Create column in metadata that is "days since first sample"
# metadata <- metadata %>% 
#   mutate(Date_collected = as.Date(Date_collected)) %>%
#   group_by(Zoo) %>%
#   arrange(Zoo, Date_collected) %>%  # ensure sorted within group
#   mutate(
#     BaselineDate = first(Date_collected),             # the earliest date in each group
#     DaysSinceBaseline = as.numeric(Date_collected - BaselineDate)  # difference in days
#   ) %>%
#   mutate(
#     # find total span per Zoo
#     TotalDays = max(DaysSinceBaseline, na.rm = TRUE),
#     Progression = DaysSinceBaseline / TotalDays
#   ) %>%
#   ungroup() %>% 
#   select(Fastq_name, Sample_ID, Animal_Name, Date_collected, Sex, Group, Specific_Group, Zoo, Tapir, Progression, Storage_reagent, `QC read count (Millions)`) %>%
#   as.data.frame()

# Now that the data has been cleaned and processed and ordered, load in the clean dataset
metadata <- read.delim(file = "../Metadata_processed.tsv", check.names = FALSE) %>% 
  mutate(Sample_ID = paste(Animal_Code, Time_Point, sep = "_")) %>%
  mutate(Date_collected = as.Date(Date_collected, format = "%m/%d/%Y"))

row.names(metadata) <- metadata$Fastq_name
```

## Get the lineage table

```{r}
lineage_table <- read.delim(file = "../Data/Lineage_table.tsv", 
                            header = TRUE, sep = "\t")
```

## Get Kraken 130 my custom DB, Domain, Phylum, Genus, Species (Species is Bracken)

```{r}
# # Domain
# # Need to make the Viruses row
# temp <- read.delim(file = "../Data/domain_abundance_table_130.tsv", 
#                    sep = "\t", 
#                    header = TRUE, 
#                    check.names = FALSE) %>% 
#   filter(str_detect(row.names(.), "(viria)|([vV]irus)|(satellites)", 
#                     negate = TRUE))
# 
# kraken_130_domain <- read.delim(file = "../Data/domain_abundance_table_130.tsv", 
#                          sep = "\t", 
#                          header = TRUE, 
#                          check.names = FALSE) %>% 
#   filter(str_detect(row.names(.),"(viria)|([vV]irus)|(satellites)")) %>% 
#   summarise_all(sum) %>% 
#   `rownames<-`(c("Viruses")) %>% 
#   rbind(temp, .) %>% 
#   select(-"86_S76") # Remove unlabled sample
# 
# rm(temp)
# 
# 
# # Phylum
kraken_130_phylum <- read.delim(file = "../Data/CombinedTableAllSamples_Bracken_Phyla.tsv", 
                                check.names = FALSE) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rownames(.) %notin% c("88_S106", "86_S76"))
# 
# 
# # Genus
# kraken_130_genus <- read.delim(file = "../Data/genus_abundance_table_130.tsv",
#                                  sep = "\t", 
#                                  header = TRUE, 
#                                  check.names = FALSE) %>% 
#   t() %>% 
#   as.data.frame() %>% 
#   select(-unclassified, -Mapped_to_upper_taxons) %>% 
#   filter(row.names(.) != "86_S76") # Remove unlabled sample


# Species, which must be treated special
kraken_130_species <- read.delim(file = "../Data/CombinedTableAllSamples_Bracken.tsv", #species_abundance_table_130
                                 sep = "\t", 
                                 header = TRUE, 
                                 check.names = FALSE, 
                                 row.names = 1) %>% 
  select(-"86_S76", -"88_S106") %>% # Remove unlabled sample and possibly contaminated wild sample
  t() %>% 
  as.data.frame()
```

## Split abundance table into Bacteria, Archaea, and Fungi and rarefy

```{r}
abund_bac <- kraken_130_species %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Bacteria"]))

abund_arch <- kraken_130_species %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Archaea"]))

abund_fun <- kraken_130_species %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Fungi"]))

abund_plant <- kraken_130_species %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Plants"]))

abund_virus <- kraken_130_species %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Viruses"]))
```

## Rarefaction groups for Mountain tapirs only

```{r}
abund_bac <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir == "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Bacteria"]))

abund_arch <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir == "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Archaea"]))

abund_fun <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir == "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Fungi"]))

abund_plant <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir == "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Plants"]))

abund_virus <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir == "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Viruses"]))
```

## Perform rarefaction

For bacteria

```{r}
# Get the min library depth
raremax <- min(rowSums(abund_bac))
# Perform rarefaction
set.seed(1025)
abund_spec_bac_rare <- rrarefy(abund_bac,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_bac_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For Archaea

```{r}
# Get the min library depth
raremax <- min(rowSums(abund_arch))
# Perform rarefaction
set.seed(1025)
abund_spec_arch_rare <- rrarefy(abund_arch,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_arch_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For Fungi

```{r}
# Get the min library depth
raremax <- min(rowSums(abund_fun))
# Perform rarefaction
set.seed(1025)
abund_spec_fun_rare <- rrarefy(abund_fun,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_fun_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For plants

```{r}
# Get the min library depth
raremax <- min(rowSums(abund_plant))
# Perform rarefaction
set.seed(1025)
abund_spec_plant_rare <- rrarefy(abund_plant,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_plant_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For viruses

```{r}
# Get the min library depth
raremax <- min(rowSums(abund_virus))
# Perform rarefaction
set.seed(1025)
abund_spec_virus_rare <- rrarefy(abund_virus,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_virus_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

#### Make the isolated domain abundance tables (Species)

```{r}
# FULL abundance matrices by domain
abund_spec_bac <- kraken_130_species %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Bacteria"]))

abund_spec_fun <- kraken_130_species %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Fungi"]))

abund_spec_arch <- kraken_130_species %>%
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Archaea"]))

# RAREFIED abundance matrices by domain
# abund_spec_bac_rare <- data_abund_rare %>% 
#   select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Bacteria"]))
# 
# abund_spec_fun_rare <- data_abund_rare %>% 
#   select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Fungi"]))
# 
# abund_spec_arch_rare <- data_abund_rare %>%
#   select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Archaea"]))
```

Now let's make sure that the metadata and the database have the samples listed in the same order

```{r}
# Check if the samples are ordered already
identical(rownames(kraken_130_species), rownames(metadata)) 
identical(rownames(abund_spec_bac), rownames(metadata))
identical(rownames(abund_spec_bac_rare), rownames(metadata))
identical(rownames(abund_spec_fun), rownames(metadata))
identical(rownames(abund_spec_fun_rare), rownames(metadata))
identical(rownames(abund_spec_arch), rownames(metadata))
identical(rownames(abund_spec_arch_rare), rownames(metadata))
identical(rownames(kraken_130_phylum), rownames(metadata))

# If they are not ordered, order them
# metadata <- metadata %>% 
#   mutate(Fastq_name = factor(Fastq_name, levels = c(rownames(kraken_130_species)))) %>% 
#   group_by(Fastq_name) %>% 
#   arrange(Fastq_name) %>% 
#   as.data.frame()
# rownames(metadata) <- metadata$Fastq_name
# 
# identical(rownames(kraken_130_species), rownames(metadata)) #TRUE
# identical(rownames(kraken_130_phylum), rownames(metadata)) #TRUE

# Save the metadata table, now that it has what we need and lists samples in the correct order
#write.table(metadata, file = "../Metadata_processed.tsv", sep = "\t", )
```

## Now I want an abundance table with nicer sample names

```{r}
# Grab the data
data_graph <- kraken_130_phylum

# Check that the metadata and data frame rows (samples) are in the same order
identical(rownames(data_graph), metadata$Fastq_name) # TRUE

# Combine the metadata$Sample_ID with the abundance table
data_graph <- cbind(metadata$Sample_ID, data_graph) %>% 
  rename(X = "metadata$Sample_ID")
rownames(data_graph) <- NULL
data_graph <- data_graph %>% 
  column_to_rownames(var = "X") %>%
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column( var = "X") %>% 
  as.data.frame()

# Now I only have to grab once
# write.table(data_graph, file = "../Data/CombinedTableAllSamples_Bracken_Phyla_Sample_ID.tsv",
#             quote = FALSE, row.names = FALSE, sep = "\t")
```
