# Boiler Plate

```{r}
library(readxl)
library(dplyr)
library(tibble)
library(vegan)
```

# Functions

```{r}
`%notin%` <- function (x, y) {!(x %in% y)}
```

# Load data sets

## Process metadata

```{r}
metadata <- read_excel(path = "../metadata_clean_deindentified.xlsx",
                       sheet = 1, 
                       col_types = c(rep("guess",2), "date", rep("guess", 7)))
row.names(metadata) <- metadata$Fastq_name

# Separate the metadata for mountain tapirs and the dam-calf-sire pairs
metadata_mt <- metadata %>%
  filter(Tapir == "Mountain")
metadata_dcs <- metadata %>% 
  filter(Tapir != "Mountain")
```

## Get the lineage table

```{r}
lineage_table <- read.delim(file = "../Data/Lineage_table.tsv", 
                            header = TRUE, sep = "\t")
```

## Get Kraken/Braken results

```{r}

kraken_130_species <- read.delim(file = "../Data/CombinedTableAllSamples_Bracken.tsv", #species_abundance_table_130
                                 sep = "\t", 
                                 header = TRUE, 
                                 check.names = FALSE, 
                                 row.names = 1) %>% 
  select(-"86_S76", -"88_S106") %>% # Remove unlabled sample and possibly contaminated wild sample
  t() %>% 
  as.data.frame()

# Check if the samples are ordered already
identical(rownames(kraken_130_species), rownames(metadata)) 
```

## Split abundance table into higher taxa and check order of samples

```{r}

abund_bac <- kraken_130_species %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Bacteria"]))
# Check order of samples against metadata
identical(rownames(abund_bac), rownames(metadata))

abund_arch <- kraken_130_species %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Archaea"]))
# Check order of samples against metadata
identical(rownames(abund_arch), rownames(metadata))

abund_fun <- kraken_130_species %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Fungi"]))
# Check order of samples against metadata
identical(rownames(abund_fun), rownames(metadata))

abund_plant <- kraken_130_species %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Plants"]))
# Check order of samples against metadata
identical(rownames(abund_plant), rownames(metadata))

abund_virus <- kraken_130_species %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Viruses"]))
# Check order of samples against metadata
identical(rownames(abund_virus), rownames(metadata))
```

# Perform rarefaction of each higher taxa for mountain tapirs only

For bacteria

```{r}
# Isolate bacterial read data for the mountain tapir samples
abund_bac_mt <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir == "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Bacteria"]))

# Get the min library depth
raremax <- min(rowSums(abund_bac_mt))
# Perform rarefaction
set.seed(1025)
abund_spec_bac_rare_mt <- rrarefy(abund_bac_mt,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_bac_rare_mt %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For Archaea

```{r}
# Isolate archaeal read data for the mountain tapir samples
abund_arch_mt <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir == "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Archaea"]))

# Get the min library depth
raremax <- min(rowSums(abund_arch_mt))
# Perform rarefaction
set.seed(1025)
abund_spec_arch_rare_mt <- rrarefy(abund_arch_mt,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_arch_rare_mt %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For Fungi

```{r}
# Isolate fungal read data for the mountain tapir samples
abund_fun_mt <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir == "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Fungi"]))

# Get the min library depth
raremax <- min(rowSums(abund_fun_mt))
# Perform rarefaction
set.seed(1025)
abund_spec_fun_rare_mt <- rrarefy(abund_fun_mt,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_fun_rare_mt %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For plants

```{r}
# Isolate plant read data for the mountain tapir samples
abund_plant_mt <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir == "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Plants"]))

# Get the min library depth
raremax <- min(rowSums(abund_plant_mt))
# Perform rarefaction
set.seed(1025)
abund_spec_plant_rare_mt <- rrarefy(abund_plant_mt,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_plant_rare_mt %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For viruses

```{r}
# Isolate viral read data for the mountain tapir samples
abund_virus_mt <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir == "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Viruses"]))

# Get the min library depth
raremax <- min(rowSums(abund_virus_mt))
# Perform rarefaction
set.seed(1025)
abund_spec_virus_rare_mt <- rrarefy(abund_virus_mt,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_virus_rare_mt %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

Check that the order of samples in the data frame and the metadata are in the same order

```{r}
identical(rownames(abund_spec_bac_rare_mt), rownames(metadata_mt))
identical(rownames(abund_spec_fun_rare_mt), rownames(metadata_mt))
identical(rownames(abund_spec_arch_rare_mt), rownames(metadata_mt))
identical(rownames(abund_spec_virus_rare_mt), rownames(metadata_mt))
identical(rownames(abund_spec_plant_rare_mt), rownames(metadata_mt))
```

# Perform rarefaction for higher taxa using dam-calf-sire samples only

For bacteria

```{r}
# Isolate bacterial read data for the mountain tapir samples
abund_bac_dcs <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir != "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Bacteria"]))

# Get the min library depth
raremax <- min(rowSums(abund_bac_dcs))
# Perform rarefaction
set.seed(1025)
abund_spec_bac_rare_dcs <- rrarefy(abund_bac_dcs,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_bac_rare_dcs %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For Archaea

```{r}
# Isolate archaeal read data for the mountain tapir samples
abund_arch_dcs <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir != "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Archaea"]))

# Get the min library depth
raremax <- min(rowSums(abund_arch_dcs))
# Perform rarefaction
set.seed(1025)
abund_spec_arch_rare_dcs <- rrarefy(abund_arch_dcs,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_arch_rare_dcs %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For Fungi

```{r}
# Isolate fungal read data for the mountain tapir samples
abund_fun_dcs <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir != "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Fungi"]))

# Get the min library depth
raremax <- min(rowSums(abund_fun_dcs))
# Perform rarefaction
set.seed(1025)
abund_spec_fun_rare_dcs <- rrarefy(abund_fun_dcs,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_spec_fun_rare_dcs %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

Check that the order of samples in the data frame and the metadata are in the same order

```{r}
identical(rownames(abund_spec_bac_rare_dcs), rownames(metadata_dcs))
identical(rownames(abund_spec_fun_rare_dcs), rownames(metadata_dcs))
identical(rownames(abund_spec_arch_rare_dcs), rownames(metadata_dcs))
```
