# Boiler Plate

```{r}
library(readxl)
library(dplyr)
library(tibble)
library(vegan)
```

# Load data sets

## Process metadata

```{r}
metadata <- read.table(file = "../metadata_clean_deindentified.tsv",
                       header = TRUE, 
                       sep = "\t",
                       colClasses = c(rep("character",3), "Date", rep("character", 6), "numeric")) %>%
  as.data.frame() # Because maaslin2 is specific about data frames
row.names(metadata) <- metadata$Fastq_name

# Separate the metadata for mountain tapirs and the dam-calf-sire pairs
metadata_mt <- metadata %>%
  filter(Tapir == "Mountain")
# Select first samples from each individual
metadata_mt_select <- metadata %>% 
  filter(rownames(.) %in% c("1_S1", "10_S5", "19_S12", "2_S2", "31_S23", "33_S25", 
                            "35_S27", "37_S29", "48_S40", "49_S41", "6_S3",
                            "87_S105", "9_S4")) 
# Select only ex-situ mountain tapirs
metadata_mt_exsitu <- metadata_mt %>% 
  filter(Group == "Ex-situ")

# Grab the dam-calf-sire pairs
metadata_dcs <- metadata %>% 
  filter(Tapir != "Mountain")
```

## Get the lineage table

```{r}
lineage_table <- read.delim(file = "../Data/Lineage_table.tsv", 
                            header = TRUE, sep = "\t")
```

## Get Kraken/Braken results

```{r}

kraken_130_species <- read.delim(file = "../Data/CombinedTableAllSamples_Bracken.tsv", #species_abundance_table_130
                                 sep = "\t", 
                                 header = TRUE, 
                                 check.names = FALSE, 
                                 row.names = 1) %>% 
  select(-"86_S76", -"88_S106") %>% # Remove unlabled sample and possibly contaminated wild sample
  t() %>% 
  as.data.frame()

# Check if the samples are ordered already
identical(rownames(kraken_130_species), rownames(metadata)) 
```

# Perform rarefaction of each higher taxa for the two mountain tapir analyses

## Wild v. *Ex-situ* (only first samples taken are used)

For bacteria

```{r}
# Isolate bacterial read data for the mountain tapir samples
abund_bac_mt_select <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% c("1_S1", "10_S5", "19_S12", "2_S2", "31_S23",
                                             "33_S25", "35_S27", "37_S29", "48_S40",
                                             "49_S41", "6_S3", "87_S105", "9_S4")) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Bacteria"]))  %>% 
  .[, colSums(. != 0) != 0] # remove species not in any samples of the subset

# Get the min library depth
raremax <- min(rowSums(abund_bac_mt_select))
# Perform rarefaction
set.seed(1025)
abund_bac_mt_select_rare <- rrarefy(abund_bac_mt_select,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_bac_mt_select_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For Archaea

```{r}
# Isolate archaeal read data for the mountain tapir samples
abund_arch_mt_select <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% c("1_S1", "10_S5", "19_S12", "2_S2", "31_S23",
                                             "33_S25", "35_S27", "37_S29", "48_S40",
                                             "49_S41", "6_S3", "87_S105", "9_S4")) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Archaea"])) %>% 
  .[, colSums(. != 0) != 0] # remove species not in any samples of the subset

# Get the min library depth
raremax <- min(rowSums(abund_arch_mt_select))
# Perform rarefaction
set.seed(1025)
abund_arch_mt_select_rare <- rrarefy(abund_arch_mt_select,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_arch_mt_select_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For Fungi

```{r}
# Isolate fungal read data for the mountain tapir samples
abund_fun_mt_select <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% c("1_S1", "10_S5", "19_S12", "2_S2", "31_S23",
                                             "33_S25", "35_S27", "37_S29", "48_S40",
                                             "49_S41", "6_S3", "87_S105", "9_S4")) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Fungi"])) %>% 
  .[, colSums(. != 0) != 0] # remove species not in any samples of the subset

# Get the min library depth
raremax <- min(rowSums(abund_fun_mt_select))
# Perform rarefaction
set.seed(1025)
abund_fun_mt_select_rare <- rrarefy(abund_fun_mt_select,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_fun_mt_select_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For viruses (no rarefaction needed, going straight to differential abundance analysis)

```{r}
abund_virus_mt_select <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% c("1_S1", "10_S5", "19_S12", "2_S2", "31_S23",
                                             "33_S25", "35_S27", "37_S29", "48_S40",
                                             "49_S41", "6_S3", "87_S105", "9_S4")) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Viruses"])) %>% 
  .[, colSums(. != 0) != 0] # remove species not in any samples of the subset

```

For plants (no rarefaction needed, going straight to differential abundance analysis)

```{r}
abund_plant_mt_select <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% c("1_S1", "10_S5", "19_S12", "2_S2", "31_S23",
                                             "33_S25", "35_S27", "37_S29", "48_S40",
                                             "49_S41", "6_S3", "87_S105", "9_S4")) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Plants"])) %>% 
  .[, colSums(. != 0) != 0] # remove species not in any samples of the subset
```

Check that the order of samples in the data frame and the metadata are in the same order

```{r}
identical(rownames(abund_bac_mt_select_rare), rownames(metadata_mt_select))
identical(rownames(abund_arch_mt_select_rare), rownames(metadata_mt_select))
identical(rownames(abund_fun_mt_select_rare), rownames(metadata_mt_select))
identical(rownames(abund_virus_mt_select), rownames(metadata_mt_select))
identical(rownames(abund_plant_mt_select), rownames(metadata_mt_select))
```

## *Ex-situ* mountain tapirs only (across time)

For bacteria

```{r}
# Isolate bacterial read data for the mountain tapir samples
abund_bac_mt_exsitu <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% rownames(metadata_mt_exsitu)) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Bacteria"])) %>% 
  .[, colSums(. != 0) != 0] # remove species not in any samples of the subset

# Get the min library depth
raremax <- min(rowSums(abund_bac_mt_exsitu))
# Perform rarefaction
set.seed(1025)
abund_bac_mt_exsitu_rare <- rrarefy(abund_bac_mt_exsitu,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_bac_mt_exsitu_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For archaea

```{r}
# Isolate archaea read data for the mountain tapir samples
abund_arch_mt_exsitu <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% rownames(metadata_mt_exsitu)) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Archaea"])) %>% 
  .[, colSums(. != 0) != 0] # remove species not in any samples of the subset

# Get the min library depth
raremax <- min(rowSums(abund_arch_mt_exsitu))
# Perform rarefaction
set.seed(1025)
abund_arch_mt_exsitu_rare <- rrarefy(abund_arch_mt_exsitu,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_arch_mt_exsitu_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For fungi

```{r}
# Isolate fungal read data for the mountain tapir samples
abund_fun_mt_exsitu <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% rownames(metadata_mt_exsitu)) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Fungi"])) %>% 
  .[, colSums(. != 0) != 0] # remove species not in any samples of the subset

# Get the min library depth
raremax <- min(rowSums(abund_fun_mt_exsitu))
# Perform rarefaction
set.seed(1025)
abund_fun_mt_exsitu_rare <- rrarefy(abund_fun_mt_exsitu,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_fun_mt_exsitu_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

Check that the order of samples in the data frame and the metadata are in the same order

```{r}
identical(rownames(abund_bac_mt_exsitu_rare), rownames(metadata_mt_exsitu))
identical(rownames(abund_arch_mt_exsitu_rare), rownames(metadata_mt_exsitu))
identical(rownames(abund_fun_mt_exsitu_rare), rownames(metadata_mt_exsitu))
```

# Perform rarefaction for higher taxa using dam-calf-sire samples only

For bacteria

```{r}
# Isolate bacterial read data for the mountain tapir samples
abund_bac_dcs <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir != "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Bacteria"])) %>% 
  .[, colSums(. != 0) != 0] # remove species not in any samples of the subset

# Get the min library depth
raremax <- min(rowSums(abund_bac_dcs))
# Perform rarefaction
set.seed(1025)
abund_bac_dcs_rare <- rrarefy(abund_bac_dcs,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_bac_dcs_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For archaea

```{r}
# Isolate archaeal read data for the mountain tapir samples
abund_arch_dcs <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir != "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Archaea"])) %>% 
  .[, colSums(. != 0) != 0] # remove species not in any samples of the subset

# Get the min library depth
raremax <- min(rowSums(abund_arch_dcs))
# Perform rarefaction
set.seed(1025)
abund_arch_dcs_rare <- rrarefy(abund_arch_dcs,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_arch_dcs_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

For Fungi

```{r}
# Isolate fungal read data for the mountain tapir samples
abund_fun_dcs <- kraken_130_species %>% 
  filter(rownames(kraken_130_species) %in% metadata$Fastq_name[metadata$Tapir != "Mountain"]) %>% 
  select(any_of(lineage_table$Species[lineage_table$Higher_taxa == "Fungi"])) %>% 
  .[, colSums(. != 0) != 0] # remove species not in any samples of the subset

# Get the min library depth
raremax <- min(rowSums(abund_fun_dcs))
# Perform rarefaction
set.seed(1025)
abund_fun_dcs_rare <- rrarefy(abund_fun_dcs,raremax) %>%
  t() %>% 
  as.data.frame() %>% 
  filter(rowSums(across(everything())) > 0) %>%
  t() %>% 
  as.data.frame()
  
# Check that this worked-- YES
abund_fun_dcs_rare %>% 
  t() %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))
```

Check that the order of samples in the data frame and the metadata are in the same order

```{r}
identical(rownames(abund_bac_dcs_rare), rownames(metadata_dcs))
identical(rownames(abund_arch_dcs_rare), rownames(metadata_dcs))
identical(rownames(abund_fun_dcs_rare), rownames(metadata_dcs))
```
