# Boiler Plate

```{r}
library(Maaslin2) # Maaslin2()
library(vroom)
library(ggrepel)
library(ggtext)
library(microbiome) # core_members()
library(hilldiv) # tss()
```

# Function

```{r}
# Graph and save volcano plot
volcano_plot <- function(maaslin_file, output_path, output_name, cutoff = 0.05, Plot_title = "Wild v. *ex-situ*", chosen = c()) {
  
  # Load the data
  taxa <- vroom(maaslin_file, altrep = FALSE)
  
  taxa$symbol=taxa$feature # We want the feature names to be the labels.
  
  sigUp <- which(taxa$coef>0 & taxa$qval<cutoff) # right top quad
  sigDown <- which(taxa$coef<0 & taxa$qval<cutoff) # left top quad
  nonSigUp <- which(taxa$coef>0 & taxa$qval>cutoff) # right bottom quad
  nonSigDown <- which(taxa$coef<0 & taxa$qval>cutoff) # left bottom quad
  
  taxa$Significance <- "FDR<1e-2, Increased Abundance"
  taxa[ sigUp, "Significance"] <- "FDR<0.05, Increased Abundance"
  taxa[ sigDown, "Significance"] <- "FDR<0.05, Decreased Abundance"
  taxa[ nonSigUp, "Significance"] <- "FDR>0.05, Increased Abundance"
  taxa[ nonSigDown, "Significance"] <- "FDR>0.05, Decreased Abundance"
  
  
  # Make the volcano plot ------------------------------------------------------
  
  #pdf(file=paste(output_path, paste(output_name, "pdf", sep = "."), sep = "/"), height = 24, width = 24) # Open the pdf
  
  ggplot(taxa, aes(x = coef, y = -log10(qval))) +
    geom_point(aes(fill = Significance, shape = Significance, color = Significance), size = 1.75) +
    geom_hline(yintercept = -log10(0.05), color = "darkred", linetype = "dashed") + 
    scale_fill_manual(values = c("#99FF33", "#A6A6A6", "#B6EC80", "#D1D1D1")) + # "#6FDE00", "#A6A6A6"
    scale_color_manual(values = c("black", "black", "#B6EC80", "#D1D1D1")) + # #696969
    scale_shape_manual(values = c(21,21,19,19)) + 
    xlab("Effect size") +
    ylab(expression('-Log'[10]*'P-Value')) +
    ggtitle(Plot_title) +
    theme_bw() + 
    guides(fill = guide_legend(nrow = 2, ncol = 2, byrow = TRUE)) +
    theme(text = element_text(size = 12),
          axis.title = element_text(size = 10),
          legend.position = "bottom", 
          legend.text = element_text(size = 10),
          legend.title = element_blank(), # element_text(size = 12),
          #legend.key.height=unit(2, "cm"),
          plot.title= element_markdown(),
          title = element_text(size = 10),
          legend.margin = margin(-10, 25, 0, 0) #t,r,b,l
          # panel.background = element_rect(fill = "black")
    ) +
    geom_text_repel(
      data = subset(taxa, feature %in% chosen), # qval < cutoff), # Select the dots that get labels
      aes(label = symbol),
      size = 3,
      box.padding = unit(0.4, "lines"),
      point.padding = unit(0.3, "lines"),
      max.overlaps = 25,
      force = 25,   # pushes labels away from each other
      force_pull = 1, # pull labels toward points
      min.segment.length = 0  # ensures a segment is always drawn
    )
  
  
  #dev.off() # Close the pdf
  
  # # Figure PDF
  # ggsave(paste(output_path, 
  #              paste(output_name, "Figure.pdf", sep = "_"), 
  #              sep = "/"), 
  #        width = 14, 
  #        height = 7)
  # 
  # # Figure SVG
  # ggsave(paste(output_path, 
  #              paste(output_name, "svg", sep = "."), 
  #              sep = "/"), 
  #        width = 14, 
  #        height = 7)
  # 
  #   # Figure SVG
  # ggsave(paste(output_path, 
  #              paste(output_name, "png", sep = "."), 
  #              sep = "/"), 
  #        width = 14, 
  #        height = 7)
  
  return(last_plot())
}
```

# Apply abundance and prevalence filters

```{r}
# Get taxa proportions
data_to_filter <- kraken_130_species %>% # Only do selected mt samples \/
  filter(rownames(kraken_130_species) %in% c("1_S1", "10_S5", "19_S12", "2_S2", "31_S23",
                                             "33_S25", "35_S27", "37_S29", "48_S40",
                                             "49_S41", "6_S3", "87_S105", "9_S4")) %>%
  t() %>% 
  tss()

# Check that this worked (samples should sum to 1)
data_to_filter %>% 
  as.data.frame() %>% 
  summarise(across(everything(), ~ sum(., na.rm = TRUE)))

# Get the list of microbes that passed filtering thresholds
microbes_pass_filter <- core_members(data_to_filter, 
                                     detection = 0.00005, 
                                     prevalence = 0.1)

# No longer need the table of values
rm(data_to_filter)
```

# Wild v. select *Ex-situ*

## For Bacteria

```{r}
# Select the microbes that passed filtering for analysis
Mas_data <- abund_bac_mt_select %>%
  select(any_of(microbes_pass_filter))

Maaslin2(Mas_data, 
         metadata_mt_select, 
         '../Data/Maaslin2_outs_rare/No rare select min detect/Bacteria', # output location
         min_abundance = 0,
         min_prevalence = 0, # Microbes have already been filtered by prevalence and abundance
         max_significance = 0.05, #qval threshold for significance
         analysis_method = 'LM',
         correction = "BH",
         fixed_effects = c("Group"),
         standardize = FALSE, #z-scores for continuous metadata, none of the metadata in the study is continuous
         transform = 'LOG',
         normalization = 'TSS', 
         reference = c('Group,Wild'))
```

### Make and save the volcano plots

```{r}
volc_bac <- volcano_plot("../Data/Maaslin2_outs_rare/No rare select min detect/Bacteria/all_results.tsv", 
             "../Data/Maaslin2_outs_rare/No rare select min detect/Bacteria",
             "Bacteria_only", Plot_title = "Wild v. *ex-situ* (Bacteria)",
             chosen = c("Ruminococcus.albus", "Treponema.bryantii", "Treponema.ruminis",
                        "Fibrobacter.succinogenes", "Ruminococcus.champanellensis", 
                        "Escherichia.coli", "Butyricimonas.virosa", 
                        "Vibrio.parahaemolyticus", "Shigella.flexneri", 
                        "Clostridioides.difficile", "Denitrobacterium.detoxificans",
                        "Slackia.heliotrinireducens"))

# For presentation
volcano_plot("../Data/Maaslin2_outs_rare/No rare select min detect/Bacteria/all_results.tsv", 
             "../Images/Marcos",
             "Volc_Bacteria", Plot_title = "Wild v. *ex-situ* (Bacteria)",
             chosen = c("Vibrio.parahaemolyticus","Shigella.flexneri", 
                        "Escherichia.coli","Fibrobacter.succinogenes",
                        "Corynebacterium.kefirresidentii",
                        "Treponema.bryantii","Treponema.ruminis","Ruminococcus.albus",
                        "Fibrobacter.succinogenes","Ruminococcus.champanellensis",
                        "Nitratifractor.salsuginis", "Paralysiella.testudinis"))
```

Make the big figure (now that we have all the parts

```{r}
# design <- "AABBCC
# DDFFFF
# EEFFFF"

design <- "AAAABBBCCCCC
DDDDDFFFFFFF
EEEEEFFFFFFF"


(p1 + p2 + p3 + plot_bac_Group_Jac + plot_bac_Group_BC + volc_bac) + 
  plot_layout(design = design) + 
  plot_annotation(tag_levels = "A", 
                  title = "") & 
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown(), 
        legend.text = element_text(size = 8, margin = margin(b = 2)), 
        legend.title = element_blank(), 
        legend.key.spacing = unit(0.1, "cm"))



# ggsave(filename = "../Images/Manuscript (Frontiers)/rare min detect/WvE_bac_Div_Diff_panel.svg",
#        width = 180,
#        height = 178,
#        units = "mm")

```

## For Archaea

```{r}
# TSS normaliza abundance matrix, so that min_abund flag in Maaslin2 actually does min_relative_abundance
Mas_data <- abund_arch_mt_select %>%
  select(any_of(microbes_pass_filter))

Maaslin2(Mas_data, 
         metadata_mt_select, 
         '../Data/Maaslin2_outs_rare/No rare select min detect/Archaea', # output location
         min_abundance = 0,
         min_prevalence = 0, # Microbes have already been filtered by prevalence and abundance
         max_significance = 0.05, #qval threshold for significance
         analysis_method = 'LM',
         correction = "BH",
         fixed_effects = c("Group"),
         standardize = FALSE, #z-scores for continuous metadata, none of the metadata in the study is continuous
         transform = 'LOG',
         normalization = 'TSS', 
         reference = c('Group,Wild'))
```

### Make and save the volcano plots

```{r}
volc_arch <- volcano_plot("../Data/Maaslin2_outs_rare/No rare select min detect/Archaea/all_results.tsv",
             "../Data/Maaslin2_outs_rare/No rare select min detect/MountainWvC/Archaea",
             "Archaea_only", Plot_title = "Wild v. *ex-situ* (Archaea)",
             chosen = c("Methanobrevibacter.ruminantium", "Methanobrevibacter.smithii",
                        "Methanoculleus.chikugoensis", "Methanocorpusculum.labreanum",
                        "Methanoculleus.bourgensis", "Methanobrevibacter.intestini", 
                        "Methanobrevibacter.olleyae", "Methanobrevibacter.millerae",
                        "Methanobrevibacter.sp..AbM4", 
                        "Methanobrevibacter.sp..TLL.48.HuF1"))

# For presentation
volcano_plot("../Data/Maaslin2_outs_rare/No rare select min detect/Archaea/all_results.tsv",
             "../Images/Marcos",
             "Volc_Archaea", Plot_title = "Wild v. *ex-situ* (Archaea)",
             chosen = c("Methanogenium.sp..S4BF","Methanoculleus.bourgensis",
                        "Methanobrevibacter.smithii","Methanobrevibacter.intestini",
                        "Methanobrevibacter.olleyae","Methanobrevibacter.ruminantium",
                        "Methanobrevibacter.sp..TLL.48.HuF1","Methanobrevibacter.sp..AbM4"))

```

Make the big plot (now that we have all the pieces)

```{r}

# design <- "AABBCC
# DDDEEE
# FFFFFF
# FFFFFF"

design <- "AAAABBBCCCCC
DDDDDFFFFFFF
EEEEEFFFFFFF"

(p4 + p5 + p6 + plot_arch_Group_Jac + plot_arch_Group_BC + volc_arch) + 
  plot_layout(design = design) + 
  plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown(),
        legend.text = element_text(size = 8, margin = margin(b = 2)), 
        legend.title = element_blank(), 
        legend.key.spacing = unit(0.1, "cm")) 

ggsave(filename = "../Images/Manuscript (Frontiers)/rare min detect/WvE_arch_Div_Diff_panel.svg", 
       width = 180,
       height = 178,
       units = "mm")
```

## For Fungi

```{r}
Mas_data <- abund_fun_mt_select %>% 
  select(any_of(microbes_pass_filter))

Maaslin2(Mas_data, 
         metadata_mt_select, 
         '../Data/Maaslin2_outs_rare/No rare select min detect/Fungi', # output location
         min_abundance = 0,
         min_prevalence = 0, # Microbes have already been filtered by prevalence and abundance
         max_significance = 0.05, #qval threshold for significance
         analysis_method = 'LM',
         correction = "BH",
         fixed_effects = c("Group"),
         standardize = FALSE, #z-scores for continuous metadata, none of the metadata in the study is continuous
         transform = 'LOG',
         normalization = 'TSS', 
         reference = c('Group,Wild'))
```

### Make and save the volcano plots

```{r}
volc_fun <- volcano_plot("../Data/Maaslin2_outs_rare/No rare select min detect/Fungi/all_results.tsv", 
             "../Data/Maaslin2_outs_rare/No rare select min detect/Fungi",
             "Fungi_only", Plot_title = "Wild v. *ex-situ* (Fungi)")
volc_fun
```

Make the big plot (now that we have all the pieces)

```{r}

design <- "AAAABBBCCCCC
DDDDDFFFFFFF
EEEEEFFFFFFF"

(p7 + p8 + p9 + plot_fun_Group_Jac + plot_fun_Group_BC + volc_fun) + 
  plot_layout(design = design) + 
  plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown(),
        legend.text = element_text(size = 8, margin = margin(b = 2)), 
        legend.title = element_blank(), 
        legend.key.spacing = unit(0.1, "cm")) 

ggsave(filename = "../Images/Manuscript (Frontiers)/rare min detect/WvE_fun_Div_Diff_panel.png",
       width = 180,
       height = 178,
       units = "mm")

```

## For Virus

```{r}
Mas_data <- abund_virus_mt_select %>% 
  select(any_of(microbes_pass_filter))

Maaslin2(Mas_data, 
         metadata_mt_select, 
         '../Data/Maaslin2_outs_rare/No rare select min detect/Viruses', # output location
         min_abundance = 0,
         min_prevalence = 0, # Microbes have already been filtered by prevalence and abundance
         max_significance = 0.05, #qval threshold for significance
         analysis_method = 'LM',
         correction = "BH",
         fixed_effects = c("Group"),
         standardize = FALSE, #z-scores for continuous metadata, none of the metadata in the study is continuous
         transform = 'LOG',
         normalization = 'TSS', 
         reference = c('Group,Wild'))
```

### Make and save the volcano plots

```{r}
volcano_plot("../Data/Maaslin2_outs_rare/No rare select min detect/Viruses/all_results.tsv", 
             "../Data/Maaslin2_outs_rare/No rare select min detect/Viruses",
             "Viruses_only", chosen = c("Kratosvirus.quantuckense","Brigitvirus.brigit","Eponavirus.epona"))
```

## For plants

```{r}
Mas_data <- abund_plant_mt_select %>% 
  select(any_of(microbes_pass_filter))

Maaslin2(Mas_data, 
         metadata_mt_select, 
         '../Data/Maaslin2_outs_rare/No rare select min detect/Plants', # output location
         min_abundance = 0,
         min_prevalence = 0, # Microbes have already been filtered by prevalence and abundance
         max_significance = 0.05, #qval threshold for significance
         analysis_method = 'LM',
         correction = "BH",
         fixed_effects = c("Group"),
         standardize = FALSE, #z-scores for continuous metadata, none of the metadata in the study is continuous
         transform = 'LOG',
         normalization = 'TSS', 
         reference = c('Group,Wild'))
```

### Make and save the volcano plots

```{r}
volcano_plot("../Data/Maaslin2_outs_rare/No rare select min detect/Plants/all_results.tsv", 
             "../Data/Maaslin2_outs_rare/No rare select min detect/Plants",
             "Plants_only") 
```
