# Boiler Plate

```{r}
library(vegan) 
library(ggplot2)
#library(car) # Homoscedasticity test
#library(ggpubr) # Density plots
library(hilldiv) # tss()
library(ape) # pcoa
library(ggtext)
library(patchwork)
library(ggrepel)
library(openxlsx) # write.xlsx
```

# Functions

```{r}
# Removes species that are not present in any sample
# returns a beta-diversity distance matrix
beta_get <- function(data, method = "bray", binary = FALSE) {
  data_abund <- data %>% 
    t() %>% # Columns must be Samples
    as.data.frame() %>%
    filter(rowSums(.) > 0) %>% # keep rows (species) that satisfy the conditions
    t() %>% # Flip back to having samples as rows
    as.matrix() 
  
  data_beta <- vegdist(data_abund, method = method, binary = binary)
  
  return(data_beta)
}
```

# Wild v. *Ex-situ* mountain tapirs

## Calculate Beta-Diversity measures

```{r}
# Distance matrix select mountain tapirs
data_beta_Jac_bac <- beta_get(abund_bac_mt_select_rare, 
                              method = "jaccard", binary = TRUE)
data_beta_BC_bac <- beta_get(abund_bac_mt_select_rare)

data_beta_Jac_arch <- beta_get(abund_arch_mt_select_rare, 
                               method = "jaccard", binary = TRUE)
data_beta_BC_arch <- beta_get(abund_arch_mt_select_rare)

data_beta_Jac_fun <- beta_get(abund_fun_mt_select_rare, 
                              method = "jaccard", binary = TRUE)
data_beta_BC_fun <- beta_get(abund_fun_mt_select_rare)

# Check the order
identical(labels(data_beta_Jac_bac), rownames(metadata_mt_select))
identical(labels(data_beta_BC_bac), rownames(metadata_mt_select))
identical(labels(data_beta_Jac_arch), rownames(metadata_mt_select))
identical(labels(data_beta_BC_arch), rownames(metadata_mt_select))
identical(labels(data_beta_Jac_fun), rownames(metadata_mt_select))
identical(labels(data_beta_BC_fun), rownames(metadata_mt_select))
```

## Bacteria

### Calculate statistics

```{r}
# Jaccard
set.seed(1025)
adonis2(data_beta_Jac_bac ~ Group, 
        data = metadata_mt_select, 
        permutations = 10000, 
        by = "margin")

# Bray-Curtis 
set.seed(1025)
adonis2(data_beta_BC_bac ~ Group, 
        data = metadata_mt_select, 
        permutations = 10000, 
        by = "margin")

```

### Beta dispersion test for the significant values

```{r}
# Jaccard
# Calculate the values (levene test analogue)
bray_disp_Jac <- betadisper(data_beta_Jac_bac, 
                        metadata_mt_select$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_Jac, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_Jac, main = "Jaccard") 


# Bray Curtis
# Calculate the values (levene test analogue)
bray_disp_BC <- betadisper(data_beta_BC_bac, 
                        metadata_mt_select$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_BC, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_BC, main = "Bray Curtis")

```

### Plot

```{r}
# Jaccard
bray_ordination <- pcoa(data_beta_Jac_bac)
points <- bray_ordination$vectors
points2 <- cbind(points, metadata_mt_select)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 2) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  ggtitle("Jaccard") + 
  scale_fill_manual(
    name = "Group",
    breaks = c("Wild", "Ex-situ"),
    values = c(
      "Wild" = "#99FF33",
      "Ex-situ" = "#A6A6A6"
    ),
    labels = c(
      "Wild" = "Wild",
      "Ex-situ" = expression(italic("Ex-situ")))) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12))
  
plot_bray_Group_Jac


# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_bac)
points <- bray_ordination$vectors
points2 <- cbind(points, metadata_mt_select)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 2) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  ggtitle("Bray-Curtis") + 
  scale_fill_manual(
    name = "Group",
    breaks = c("Wild", "Ex-situ"),
    values = c(
      "Wild" = "#99FF33",
      "Ex-situ" = "#A6A6A6"
    ),
    labels = c(
      "Wild" = "Wild",
      "Ex-situ" = expression(italic("Ex-situ")))) +
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12))
  
plot_bray_Group_BC


(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") + 
  plot_layout(guides = "collect") &
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02),
        legend.position = "bottom") 



# Save the plot
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_bac_species_beta_div_rare.pdf",
#        height = 4,
#        width = 6,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_bac_species_beta_div_rare.tiff",
#        height = 4,
#        width = 6,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_bac_species_beta_div_rare.png",
#        height = 4,
#        width = 6,
#        units = "in")
```

## Archaea

### Calculate statistics

```{r}
# Jaccard 
set.seed(1025)
adonis2(data_beta_Jac_arch ~ Group, 
        data = metadata_mt_select, 
        permutations = 10000, 
        by = "margin")


# Bray-Curtis 
set.seed(1025)
adonis2(data_beta_BC_arch ~ Group, 
        data = metadata_mt_select, 
        permutations = 10000, 
        by = "margin")
```

### Beta dispersion test for the significant values

```{r}
# Jaccard
# Calculate the values (levene test analogue)
bray_disp_Jac <- betadisper(data_beta_Jac_arch, 
                        metadata_mt_select$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_Jac, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_Jac, main = "Jaccard") 


# Bray Curtis
# Calculate the values (levene test analogue)
bray_disp_BC <- betadisper(data_beta_BC_arch, 
                        metadata_mt_select$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_BC, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_BC, main = "Bray Curtis")

```

### Plot

```{r}
# Jaccard
bray_ordination <- pcoa(data_beta_Jac_arch)
points <- bray_ordination$vectors
points2 <- cbind(points, metadata_mt_select)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 2) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  ggtitle("Jaccard") + 
  scale_fill_manual(
    name = "Group",
    breaks = c("Wild", "Ex-situ"),
    values = c(
      "Wild" = "#99FF33",
      "Ex-situ" = "#A6A6A6"
    ),
    labels = c(
      "Wild" = "Wild",
      "Ex-situ" = expression(italic("Ex-situ")))) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12))
  
plot_bray_Group_Jac



# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_arch)
points <- bray_ordination$vectors
points2 <- cbind(points, metadata_mt_select)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 2) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  ggtitle("Bray-Curtis") + 
  scale_fill_manual(
    name = "Group",
    breaks = c("Wild", "Ex-situ"),
    values = c(
      "Wild" = "#99FF33",
      "Ex-situ" = "#A6A6A6"
    ),
    labels = c(
      "Wild" = "Wild",
      "Ex-situ" = expression(italic("Ex-situ")))) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12))
  
plot_bray_Group_BC



(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") + 
  plot_layout(guides = "collect") &
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02),
        legend.position = "bottom") 


# Save the plot
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_arch_species_beta_div_rare.pdf",
#        height = 4,
#        width = 6,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_arch_species_beta_div_rare.tiff",
#        height = 4,
#        width = 6,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_arch_species_beta_div_rare.png",
#        height = 4,
#        width = 6,
#        units = "in")
```

## Fungi

### Calculate statistics

```{r}
# Jaccard 
set.seed(1025)
adonis2(data_beta_Jac_fun ~ Group, 
        data = metadata_mt_select, 
        permutations = 10000, 
        by = "margin")


# Bray-Curtis
set.seed(1025)
adonis2(data_beta_BC_fun ~ Group, 
        data = metadata_mt_select, 
        permutations = 10000, 
        by = "margin")

```

### Beta dispersion test for the significant values

```{r}
# Jaccard
# Calculate the values (levene test analogue)
bray_disp_Jac <- betadisper(data_beta_Jac_fun, 
                        metadata_mt_select$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_Jac, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_Jac, main = "Jaccard") 


# Bray Curtis
# Calculate the values (levene test analogue)
bray_disp_BC <- betadisper(data_beta_BC_fun, 
                        metadata_mt_select$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_BC, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_BC, main = "Bray Curtis")

```

### Plot

```{r}
# Jaccard
bray_ordination <- pcoa(data_beta_Jac_fun)
points <- bray_ordination$vectors
points2 <- cbind(points, metadata_mt_select)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 2) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  ggtitle("Jaccard") + 
  scale_fill_manual(
    name = "Group",
    breaks = c("Wild", "Ex-situ"),
    values = c(
      "Wild" = "#99FF33",
      "Ex-situ" = "#A6A6A6"
    ),
    labels = c(
      "Wild" = "Wild",
      "Ex-situ" = expression(italic("Ex-situ")))) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12))
  
plot_bray_Group_Jac




# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_fun)
points <- bray_ordination$vectors
points2 <- cbind(points, metadata_mt_select)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 2) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  ggtitle("Bray-Curtis") + 
  scale_fill_manual(
    name = "Group",
    breaks = c("Wild", "Ex-situ"),
    values = c(
      "Wild" = "#99FF33",
      "Ex-situ" = "#A6A6A6"
    ),
    labels = c(
      "Wild" = "Wild",
      "Ex-situ" = expression(italic("Ex-situ")))) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12))
  
plot_bray_Group_BC




(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") + 
  plot_layout(guides = "collect") &
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02),
        legend.position = "bottom") 


# p = 0.0006999, R2 = 0.3113
# Save the plot
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_fun_species_beta_div_rare.pdf",
#        height = 4,
#        width = 6,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_fun_species_beta_div_rare.tiff",
#        height = 4,
#        width = 6,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_fun_species_beta_div_rare.png",
#        height = 4,
#        width = 6,
#        units = "in")
```

# Malayan Calf samples over time

### Set–up variables and data

```{r}
# Grab the Malayan samples and change the date to something I can graph
meta_test <- metadata_dcs %>% 
  filter(Tapir %in% c("Malayan")) %>% 
  mutate(Date_collected = as.Date(.$Date_collected, format = "%m/%d/%Y")) %>% 
  mutate(months_since_start = (time_length(interval(min(Date_collected), Date_collected), unit = "months") + 0.76))


# The abundance tables for Malayan tapirs
abund_bac_Mal_rare <- abund_bac_dcs_rare[row.names(abund_bac_dcs_rare) %in% meta_test$Fastq_name, ]

abund_arch_Mal_rare <- abund_arch_dcs_rare[row.names(abund_arch_dcs_rare) %in% meta_test$Fastq_name, ]

abund_fun_Mal_rare <- abund_fun_dcs_rare[row.names(abund_fun_dcs_rare) %in% meta_test$Fastq_name, ] 


# Distance matrices
data_beta_BC_bac_Mal <- beta_get(abund_bac_Mal_rare)
data_beta_Jac_bac_Mal <- beta_get(abund_bac_Mal_rare, method = "jaccard", binary = TRUE)

data_beta_BC_arch_Mal <- beta_get(abund_arch_Mal_rare)
data_beta_Jac_arch_Mal <- beta_get(abund_arch_Mal_rare, method = "jaccard", binary = TRUE)

data_beta_BC_fun_Mal <- beta_get(abund_fun_Mal_rare)
data_beta_Jac_fun_Mal <- beta_get(abund_fun_Mal_rare, method = "jaccard", binary = TRUE)
```

## Plot

### Bacteria

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_bac_Mal)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
  # geom_smooth(method = "auto", color = "red", se = FALSE) + 
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  scale_x_continuous(breaks = seq(0,10, by = 1), minor_breaks = NULL) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Bray-Curtis") + 
  scale_color_manual(values = c("#00bfc4","#ff64b0", "black")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_BC

# ggsave(file = "../Images/MATC_Beta_time.png", 
#         units = "in",
#         height = 3.6,
#         width = 6)
# ggsave(file = "../Images/MATC_Beta_time.svg", 
#         units = "in",
#         height = 3.6,
#         width = 6)

# Jaccard
bray_ordination <- pcoa(data_beta_Jac_bac_Mal)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  scale_x_continuous(breaks = seq(0,10, by = 1), minor_breaks = NULL) + 
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  scale_color_manual(values = c("#00bfc4","#ff64b0", "black")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        legend.position = "none",
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_Jac

(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 
```

### Archaea

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_arch_Mal)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
  # geom_smooth(method = "auto", color = "red", se = FALSE) + 
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  scale_x_continuous(breaks = seq(0,10, by = 1), minor_breaks = NULL) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Bray-Curtis") + 
  scale_color_manual(values = c("#00bfc4","#ff64b0", "black")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_BC

# ggsave(file = "../Images/MATC_Beta_time.png", 
#         units = "in",
#         height = 3.6,
#         width = 6)
# ggsave(file = "../Images/MATC_Beta_time.svg", 
#         units = "in",
#         height = 3.6,
#         width = 6)

# Jaccard
bray_ordination <- pcoa(data_beta_Jac_arch_Mal)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  scale_x_continuous(breaks = seq(0,10, by = 1), minor_breaks = NULL) + 
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  scale_color_manual(values = c("#00bfc4","#ff64b0", "black")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        legend.position = "none",
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_Jac

(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 
```

### Fungi

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_fun_Mal)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
  # geom_smooth(method = "auto", color = "red", se = FALSE) + 
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  scale_x_continuous(breaks = seq(0,10, by = 1), minor_breaks = NULL) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Bray-Curtis") + 
  scale_color_manual(values = c("#00bfc4","#ff64b0", "black")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_BC

# ggsave(file = "../Images/MATC_Beta_time.png", 
#         units = "in",
#         height = 3.6,
#         width = 6)
# ggsave(file = "../Images/MATC_Beta_time.svg", 
#         units = "in",
#         height = 3.6,
#         width = 6)

# Jaccard
bray_ordination <- pcoa(data_beta_Jac_fun_Mal)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  scale_x_continuous(breaks = seq(0,10, by = 1), minor_breaks = NULL) + 
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  scale_color_manual(values = c("#00bfc4","#ff64b0", "black")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        legend.position = "none",
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_Jac

(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 
```

# Lowland calf over time

### Set-up variables and data

```{r}
# Grab the lowland samples and change the date to something I can graph
meta_test <- metadata_dcs %>% 
  filter(Tapir %in% c("Lowland")) %>% 
  mutate(Date_collected = as.Date(.$Date_collected, format = "%m/%d/%Y")) %>% 
  mutate(months_since_start = (time_length(interval(min(Date_collected), Date_collected), unit = "months") + 0.613))

# The abundance tables
abund_bac_low_rare <- abund_bac_dcs_rare[row.names(abund_bac_dcs_rare) %in% meta_test$Fastq_name, ] 

abund_arch_low_rare <- abund_arch_dcs_rare[row.names(abund_arch_dcs_rare) %in% meta_test$Fastq_name, ] 

abund_fun_low_rare <- abund_fun_dcs_rare[row.names(abund_fun_dcs_rare) %in% meta_test$Fastq_name, ] 

# Distance matrices
data_beta_BC_bac_low <- beta_get(abund_bac_low_rare)
data_beta_Jac_bac_low <- beta_get(abund_bac_low_rare, method = "jaccard", binary = TRUE)

data_beta_BC_arch_low <- beta_get(abund_arch_low_rare)
data_beta_Jac_arch_low <- beta_get(abund_arch_low_rare, method = "jaccard", binary = TRUE)

data_beta_BC_fun_low <- beta_get(abund_fun_low_rare)
data_beta_Jac_fun_low <- beta_get(abund_fun_low_rare, method = "jaccard", binary = TRUE)
```

## Plot

### bacteria

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_bac_low)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
  # geom_smooth(method = "auto", color = "red", se = FALSE) + 
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Bray-Curtis") + 
  scale_color_manual(values = c("#c77cff","#00c08b")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_BC

# ggsave(file = "../Images/LTC_Beta_time.png", 
#         units = "in",
#         height = 3.6,
#         width = 6)
# ggsave(file = "../Images/LTC_Beta_time.svg", 
#         units = "in",
#         height = 3.6,
#         width = 6)

# Jaccard
bray_ordination <- pcoa(data_beta_Jac_bac_low)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  scale_color_manual(values = c("#c77cff","#00c08b")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        legend.position = "none",
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_Jac

(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 
```

### Archaea

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_arch_low)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
  # geom_smooth(method = "auto", color = "red", se = FALSE) + 
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Bray-Curtis") + 
  scale_color_manual(values = c("#c77cff","#00c08b")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_BC

# ggsave(file = "../Images/LTC_Beta_time.png", 
#         units = "in",
#         height = 3.6,
#         width = 6)
# ggsave(file = "../Images/LTC_Beta_time.svg", 
#         units = "in",
#         height = 3.6,
#         width = 6)

# Jaccard
bray_ordination <- pcoa(data_beta_Jac_arch_low)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  scale_color_manual(values = c("#c77cff","#00c08b")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        legend.position = "none",
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_Jac

(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 
```

### Fungi

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_fun_low)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
  # geom_smooth(method = "auto", color = "red", se = FALSE) + 
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Bray-Curtis") + 
  scale_color_manual(values = c("#c77cff","#00c08b")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_BC

# ggsave(file = "../Images/LTC_Beta_time.png", 
#         units = "in",
#         height = 3.6,
#         width = 6)
# ggsave(file = "../Images/LTC_Beta_time.svg", 
#         units = "in",
#         height = 3.6,
#         width = 6)

# Jaccard
bray_ordination <- pcoa(data_beta_Jac_fun_low)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  scale_color_manual(values = c("#c77cff","#00c08b")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        legend.position = "none",
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_Jac

(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 
```
