# Boiler Plate

```{r}
library(vegan) 
library(ggplot2)
#library(car) # Homoscedasticity test
#library(ggpubr) # Density plots
library(hilldiv) # tss()
library(ape) # pcoa
library(ggtext)
library(patchwork)
library(ggrepel)
```

# Functions

```{r}
# Returns a data frame with only specific samples based on metadata values
select_samples <- function(data, metadata, group = "", values = c()) {
  hold <- metadata$Fastq_name[metadata[[group]] %in% values]
  data_subset <- data %>% 
    t() %>% 
    as.data.frame() %>%
    select(any_of(hold)) %>% 
    t()
  
  return(data_subset)
}

# Does a TSS normalization on a data frame and then returns a beta-diversity 
# distance matrix
beta_get <- function(data, method = "bray", binary = FALSE) {
  data_abund <- data %>% 
    t() %>% # Columns must be Samples
    as.data.frame() %>%
    filter(rowSums(.) > 0) %>% # keep rows (species) that satisfy the conditions
    # tss() %>% # cols = samples 
    t() %>%
    as.matrix() 
  
  data_beta <- vegdist(data_abund, method = method, binary = binary)
  
  return(data_beta)
}
```

# Captive v. Wild

## Set-up Variables and Data

```{r}
# Metadata 
meta_test <- metadata %>% 
  filter(Specific_Group %in% c("Wild","Ex-situ")) %>% 
  group_by(Animal_Code) %>% 
  arrange(Date_collected, .by_group = TRUE) %>% 
  slice(1) %>% # grab the first date samples were taken from each animal 
  ungroup() %>% 
  mutate(Fastq_name = factor(.$Fastq_name, levels = c("1_S1", "10_S5", "19_S12", "2_S2", "31_S23", "33_S25", 
                                                      "35_S27", "37_S29", "48_S40", "49_S41", "6_S3",
                                                      "87_S105", "9_S4"))) %>% 
  arrange(Fastq_name) %>% 
  as.data.frame()
row.names(meta_test) <- meta_test$Fastq_name


# The abundance tables
Selected_bac <- abund_spec_bac_rare[row.names(abund_spec_bac_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set
Selected_arch <- abund_spec_arch_rare[row.names(abund_spec_arch_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set
Selected_fun <- abund_spec_fun_rare[row.names(abund_spec_fun_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set
Selected_virus <- abund_spec_virus_rare[row.names(abund_spec_virus_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set


# Distance matrices
data_beta_BC_bac <- beta_get(Selected_bac)
data_beta_Jac_bac <- beta_get(Selected_bac, method = "jaccard", binary = TRUE)

data_beta_BC_arch <- beta_get(Selected_arch)
data_beta_Jac_arch <- beta_get(Selected_arch, method = "jaccard", binary = TRUE)

data_beta_BC_fun <- beta_get(Selected_fun)
data_beta_Jac_fun <- beta_get(Selected_fun, method = "jaccard", binary = TRUE)

data_beta_BC_virus <- beta_get(Selected_virus)
data_beta_Jac_virus <- beta_get(Selected_virus, method = "jaccard", binary = TRUE)
```

## Bacteria

### Calculate statistics

```{r}
# Bray-Curtis P = 0.0006999
set.seed(1025)
adonis2(data_beta_BC_bac ~ Group, 
        data = meta_test, 
        permutations = 10000, 
        by = "margin")


# Jaccard P = 0.0006999
set.seed(1025)
adonis2(data_beta_Jac_bac ~ Group, 
        data = meta_test, 
        permutations = 10000, 
        by = "margin")

```

### Beta dispersion test for the significant values

```{r}
# Bray Curtis
# Calculate the values (levene test analogue)
bray_disp_BC <- betadisper(data_beta_BC_bac, 
                        meta_test$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_BC, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_BC, main = "Bray Curtis (p = 0.0466)")


# Jaccard
# Calculate the values (levene test analogue)
bray_disp_Jac <- betadisper(data_beta_Jac_bac, 
                        meta_test$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_Jac, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_Jac, main = "Jaccard (p = 0.008199)") 
```

### Plot

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_bac)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 2) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  ggtitle("Bray-Curtis") + 
  scale_fill_manual(
    name = "Group",
    breaks = c("Wild", "Ex-situ"),
    values = c(
      "Wild" = "#99FF33",
      "Ex-situ" = "#A6A6A6"
    ),
    labels = c(
      "Wild" = "Wild",
      "Ex-situ" = expression(italic("Ex-situ")))) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12))
  
plot_bray_Group_BC


# Jaccard
bray_ordination <- pcoa(data_beta_Jac_bac)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 2) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  ggtitle("Jaccard") + 
  scale_fill_manual(
    name = "Group",
    breaks = c("Wild", "Ex-situ"),
    values = c(
      "Wild" = "#99FF33",
      "Ex-situ" = "#A6A6A6"
    ),
    labels = c(
      "Wild" = "Wild",
      "Ex-situ" = expression(italic("Ex-situ")))) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12))
  
plot_bray_Group_Jac

(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") + 
  plot_layout(guides = "collect") &
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02),
        legend.position = "bottom") 


# p = 0.0006999, R2 = 0.3113
# Save the plot
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_bac_species_beta_div_rare.pdf",
#        height = 4,
#        width = 6,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_bac_species_beta_div_rare.tiff",
#        height = 4,
#        width = 6,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_bac_species_beta_div_rare.png",
#        height = 4,
#        width = 6,
#        units = "in")
```

plot for the presentation

```{r}
bray_ordination <- pcoa(data_beta_BC_bac)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 3) +
  theme_bw() + 
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Bray-Curtis (p = 0.000699*)") + 
  scale_fill_manual(values = c("#A6A6A6","#99FF33")) + 
  theme(title = element_text(size = 16),
        axis.title = element_text(size = 16), 
        axis.text = element_text(size = 14), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 13),
        legend.position = "none",
        plot.title = element_markdown()) 
  
plot_bray_Group_BC

ggsave( file = "../Images/Presentation/Beta_bac_BC.png", 
        units = "in",
        height = 3.48,
        width = 4)
```

## Archaea

### Calculate statistics

```{r}
# Bray-Curtis P = 0.0011
set.seed(1025)
adonis2(data_beta_BC_arch ~ Group, 
        data = meta_test, 
        permutations = 10000, 
        by = "margin")


# Jaccard P = 0.0006999
set.seed(1025)
adonis2(data_beta_Jac_arch ~ Group, 
        data = meta_test, 
        permutations = 10000, 
        by = "margin")

```

### Beta dispersion test for the significant values

```{r}
# Bray Curtis
# Calculate the values (levene test analogue)
bray_disp_BC <- betadisper(data_beta_BC_arch, 
                        meta_test$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_BC, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_BC, main = "Bray Curtis (p = 0.07589)")


# Jaccard
# Calculate the values (levene test analogue)
bray_disp_Jac <- betadisper(data_beta_Jac_arch, 
                        meta_test$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_Jac, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_Jac, main = "Jaccard (p = 9.999e-05)") 
```

### Plot

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_arch)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 2) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  ggtitle("Bray-Curtis") + 
  scale_fill_manual(
    name = "Group",
    breaks = c("Wild", "Ex-situ"),
    values = c(
      "Wild" = "#99FF33",
      "Ex-situ" = "#A6A6A6"
    ),
    labels = c(
      "Wild" = "Wild",
      "Ex-situ" = expression(italic("Ex-situ")))) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12))
  
plot_bray_Group_BC


# Jaccard
bray_ordination <- pcoa(data_beta_Jac_arch)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 2) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  ggtitle("Jaccard") + 
  scale_fill_manual(
    name = "Group",
    breaks = c("Wild", "Ex-situ"),
    values = c(
      "Wild" = "#99FF33",
      "Ex-situ" = "#A6A6A6"
    ),
    labels = c(
      "Wild" = "Wild",
      "Ex-situ" = expression(italic("Ex-situ")))) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12))
  
plot_bray_Group_Jac

(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") + 
  plot_layout(guides = "collect") &
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02),
        legend.position = "bottom") 

# p = 0.0006999, R2 = 0.3113
# Save the plot
ggsave(file = "../Images/Manuscript (Frontiers)/WvE_arch_species_beta_div_rare.pdf",
       height = 4,
       width = 6,
       units = "in")

ggsave(file = "../Images/Manuscript (Frontiers)/WvE_arch_species_beta_div_rare.tiff",
       height = 4,
       width = 6,
       units = "in")

ggsave(file = "../Images/Manuscript (Frontiers)/WvE_arch_species_beta_div_rare.png",
       height = 4,
       width = 6,
       units = "in")
```

Plot for the presentation

```{r}
bray_ordination <- pcoa(data_beta_BC_arch)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 3) +
  theme_bw() + 
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Bray-Curtis (p = 0.000699*)") + 
  scale_fill_manual(values = c("#A6A6A6","#99FF33")) + 
  theme(title = element_text(size = 16),
        axis.title = element_text(size = 16), 
        axis.text = element_text(size = 14), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 13),
        legend.position = "none",
        plot.title = element_markdown()) 
  
plot_bray_Group_BC

ggsave( file = "../Images/Presentation/Beta_arch_BC.png", 
        units = "in",
        height = 3.48,
        width = 4)
```

## Fungi

### Calculate statistics

```{r}
# Bray-Curtis P = 0.0281
set.seed(1025)
adonis2(data_beta_BC_fun ~ Group, 
        data = meta_test, 
        permutations = 10000, 
        by = "margin")


# Jaccard P = 0.6365
set.seed(1025)
adonis2(data_beta_Jac_fun ~ Group, 
        data = meta_test, 
        permutations = 10000, 
        by = "margin")

```

### Beta dispersion test for the significant values

```{r}
# Bray Curtis
# Calculate the values (levene test analogue)
bray_disp_BC <- betadisper(data_beta_BC_fun, 
                        meta_test$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_BC, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_BC, main = "Bray Curtis (p = 0.006699)")


# Jaccard
# Calculate the values (levene test analogue)
bray_disp_Jac <- betadisper(data_beta_Jac_fun, 
                        meta_test$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_Jac, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_Jac, main = "Jaccard (p = 0.2966)") 
```

### Plot

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_fun)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 2) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  ggtitle("Bray-Curtis") + 
  scale_fill_manual(
    name = "Group",
    breaks = c("Wild", "Ex-situ"),
    values = c(
      "Wild" = "#99FF33",
      "Ex-situ" = "#A6A6A6"
    ),
    labels = c(
      "Wild" = "Wild",
      "Ex-situ" = expression(italic("Ex-situ")))) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12))
  
plot_bray_Group_BC


# Jaccard
bray_ordination <- pcoa(data_beta_Jac_fun)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 2) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  ggtitle("Jaccard") + 
  scale_fill_manual(
    name = "Group",
    breaks = c("Wild", "Ex-situ"),
    values = c(
      "Wild" = "#99FF33",
      "Ex-situ" = "#A6A6A6"
    ),
    labels = c(
      "Wild" = "Wild",
      "Ex-situ" = expression(italic("Ex-situ")))) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12))
  
plot_bray_Group_Jac

(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") + 
  plot_layout(guides = "collect") &
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02),
        legend.position = "bottom") 


# p = 0.0006999, R2 = 0.3113
# Save the plot
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_fun_species_beta_div_rare.pdf",
#        height = 4,
#        width = 6,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_fun_species_beta_div_rare.tiff",
#        height = 4,
#        width = 6,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/WvE_fun_species_beta_div_rare.png",
#        height = 4,
#        width = 6,
#        units = "in")
```

plot for presentation

```{r}
bray_ordination <- pcoa(data_beta_BC_fun)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  fill = Group)) + 
  geom_point(shape = 21, size = 3) +
  theme_bw() + 
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  labs(fill = "Status") + 
  ggtitle("Bray-Curtis (p = 0.058)") + 
  scale_fill_manual(values = c("#A6A6A6","#99FF33")) + 
  theme(title = element_text(size = 16),
        axis.title = element_text(size = 16), 
        axis.text = element_text(size = 14), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 13),
        legend.position = "right",
        plot.title = element_markdown()) 
  
plot_bray_Group_BC

ggsave( file = "../Images/Presentation/Beta_fun_BC.png", 
        units = "in",
        height = 3.48,
        width = 5)
```

## Virus

### Calculate Statistics

```{r}
# Bray-Curtis 
set.seed(1025)
adonis2(data_beta_BC_virus ~ Group, 
        data = meta_test, 
        permutations = 10000, 
        by = "margin")


# Jaccard 
set.seed(1025)
adonis2(data_beta_Jac_virus ~ Group, 
        data = meta_test, 
        permutations = 10000, 
        by = "margin")
```

### Beta dispersion test for the significant values

```{r}
# Bray Curtis
# Calculate the values (levene test analogue)
bray_disp_BC <- betadisper(data_beta_BC_virus, 
                        meta_test$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_BC, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_BC, main = "Bray Curtis")


# Jaccard
# Calculate the values (levene test analogue)
bray_disp_Jac <- betadisper(data_beta_Jac_virus, 
                        meta_test$Group)
# Actually give me a p-value to assess results
set.seed(1025)
permutest(bray_disp_Jac, 
          permutations = how(nperm=10000))
# Show me a plot
plot(bray_disp_Jac, main = "Jaccard") 
```

### Plot

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_virus)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  color = Group)) + 
  geom_point(size = 3) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Bray-Curtis") + 
  scale_color_manual(values = c("blue4","orange2")) + 
  theme(title = element_text(size = 18),
        axis.title = element_text(size = 16), 
        axis.text = element_text(size = 14), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 13),
        plot.title = element_markdown())
  
plot_bray_Group_BC


# Jaccard
bray_ordination <- pcoa(data_beta_Jac_virus)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
set.seed(10)
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  color = Group)) + 
  geom_point(size = 3) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  scale_color_manual(values = c("blue4","orange2")) + 
  theme(title = element_text(size = 18),
        axis.title = element_text(size = 16), 
        axis.text = element_text(size = 14), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 13),
        legend.position = "none",
        plot.title = element_markdown())
  
plot_bray_Group_Jac


(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "a", 
                             title = "Wild v. *ex-situ* mt. tapirs (Virus)") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 
```

# Malayan Calf samples over time

### Set–up variables and data

```{r}
# Grab the fastq_file names
meta_test <- metadata %>% 
  filter(Tapir %in% c("Malayan")) %>% 
  mutate(Date_collected = as.Date(.$Date_collected, format = "%m/%d/%Y")) %>% 
  mutate(months_since_start = (time_length(interval(min(Date_collected), Date_collected), unit = "months") + 0.76))

# The abundance tables
Selected_bac <- abund_spec_bac_rare[row.names(abund_spec_bac_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set
Selected_arch <- abund_spec_arch_rare[row.names(abund_spec_arch_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set
Selected_fun <- abund_spec_fun_rare[row.names(abund_spec_fun_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set

# Distance matrices
data_beta_BC_bac <- beta_get(Selected_bac)
data_beta_Jac_bac <- beta_get(Selected_bac, method = "jaccard", binary = TRUE)

data_beta_BC_arch <- beta_get(Selected_arch)
data_beta_Jac_arch <- beta_get(Selected_arch, method = "jaccard", binary = TRUE)

data_beta_BC_fun <- beta_get(Selected_fun)
data_beta_Jac_fun <- beta_get(Selected_fun, method = "jaccard", binary = TRUE)
```

## Plot

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_bac)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
  # geom_smooth(method = "auto", color = "red", se = FALSE) + 
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  scale_x_continuous(breaks = seq(0,10, by = 1), minor_breaks = NULL) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Bray-Curtis") + 
  scale_color_manual(values = c("#00bfc4","#ff64b0", "black")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_BC

ggsave(file = "../Images/MATC_Beta_time.png", 
        units = "in",
        height = 3.6,
        width = 6)
ggsave(file = "../Images/MATC_Beta_time.svg", 
        units = "in",
        height = 3.6,
        width = 6)

# Jaccard
bray_ordination <- pcoa(data_beta_Jac_bac)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 3) +
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  scale_x_continuous(breaks = seq(0,10, by = 1), minor_breaks = NULL) + 
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  scale_color_manual(values = c("#00bfc4","#ff64b0", "black")) + 
  theme(title = element_text(size = 18),
        axis.title = element_text(size = 16), 
        axis.text = element_text(size = 14), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 13),
        legend.position = "none",
        plot.title = element_markdown())
  
plot_bray_Group_Jac

(plot_bray_Group_Jac | plot_bray_Group_BC) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 
```

# Lowland calf over time

### Set-up variables and data

```{r}
# Grab the fastq_file names
meta_test <- metadata %>% 
  filter(Tapir %in% c("Lowland")) %>% 
  mutate(Date_collected = as.Date(.$Date_collected, format = "%m/%d/%Y")) %>% 
  mutate(months_since_start = (time_length(interval(min(Date_collected), Date_collected), unit = "months") + 1))

# The abundance tables
Selected_bac <- abund_spec_bac_rare[row.names(abund_spec_bac_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set
Selected_arch <- abund_spec_arch_rare[row.names(abund_spec_arch_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set
Selected_fun <- abund_spec_fun_rare[row.names(abund_spec_fun_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set

# Distance matrices
data_beta_BC_bac <- beta_get(Selected_bac)
data_beta_Jac_bac <- beta_get(Selected_bac, method = "jaccard", binary = TRUE)

data_beta_BC_arch <- beta_get(Selected_arch)
data_beta_Jac_arch <- beta_get(Selected_arch, method = "jaccard", binary = TRUE)

data_beta_BC_fun <- beta_get(Selected_fun)
data_beta_Jac_fun <- beta_get(Selected_fun, method = "jaccard", binary = TRUE)
```

## Plot

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_bac)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 2) +
  # geom_smooth(method = "auto", color = "red", se = FALSE) + 
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Bray-Curtis") + 
  scale_color_manual(values = c("#c77cff","#00c08b")) + 
  theme(title = element_text(size = 12),
        axis.title = element_text(size = 12), 
        axis.text = element_text(size = 10), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 12),
        plot.title = element_markdown()) + 
  xlab("Calf age (month of life)")
  
plot_bray_Group_BC

ggsave(file = "../Images/LTC_Beta_time.png", 
        units = "in",
        height = 3.6,
        width = 6)
ggsave(file = "../Images/LTC_Beta_time.svg", 
        units = "in",
        height = 3.6,
        width = 6)

# Jaccard
bray_ordination <- pcoa(data_beta_Jac_bac)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = months_since_start, 
                                  y = Axis.1, 
                                  color = Animal_Code)) + 
  geom_point(size = 3) +
#  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  scale_color_manual(values = c("#c77cff","#00c08b")) + 
  theme(title = element_text(size = 18),
        axis.title = element_text(size = 16), 
        axis.text = element_text(size = 14), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 13),
        legend.position = "none",
        plot.title = element_markdown())
  
plot_bray_Group_Jac
```

# Beta diversity of all Animals

### Set-up variables and data

```{r}
# 
meta_test <- metadata %>% 
  filter(Sample_ID %in% c("WMT1_0","WMT2_0","WMT3_0","WMT4_0","WMT5_0","WMT6_0","LTC_27","MaTC_17","ZMT1_0","ZMT2_0","ZMT3_0","ZMT4_0","ZMT5_0","ZMT6_0","ZMT7_0","LTD_0","MaTD_0","MaTS_0"))
# The abundance tables
Selected_bac <- abund_spec_bac_rare[row.names(abund_spec_bac_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set
Selected_arch <- abund_spec_arch_rare[row.names(abund_spec_arch_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set
Selected_fun <- abund_spec_fun_rare[row.names(abund_spec_fun_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set

# Distance matrices
data_beta_BC_bac <- beta_get(Selected_bac)
data_beta_Jac_bac <- beta_get(Selected_bac, method = "jaccard", binary = TRUE)

data_beta_BC_arch <- beta_get(Selected_arch)
data_beta_Jac_arch <- beta_get(Selected_arch, method = "jaccard", binary = TRUE)

data_beta_BC_fun <- beta_get(Selected_fun)
data_beta_Jac_fun <- beta_get(Selected_fun, method = "jaccard", binary = TRUE)
```

## Plot

```{r}
# Bray Curtis
# Calculate the ordination
bray_ordination <- pcoa(data_beta_BC_bac)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_BC <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  color = Group)) + 
  geom_point(size = 3) +
  # geom_smooth(method = "auto", color = "red", se = FALSE) + 
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Bray-Curtis") + 
  # scale_color_manual(values = c("#c77cff","#00c08b")) + 
  theme(title = element_text(size = 18),
        axis.title = element_text(size = 16), 
        axis.text = element_text(size = 14), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 13),
        plot.title = element_markdown())
  
plot_bray_Group_BC

ggsave(file = "../Images/Supp_Fig_10.svg")
ggsave(file = "../Images/Supp_Fig_10.png")

# Jaccard
bray_ordination <- pcoa(data_beta_Jac_bac)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  color = Specific_Group)) + 
  geom_point(size = 3) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  # scale_color_manual(values = c("#c77cff","#00c08b")) + 
  theme(title = element_text(size = 18),
        axis.title = element_text(size = 16), 
        axis.text = element_text(size = 14), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 13),
        legend.position = "none",
        plot.title = element_markdown())
  
plot_bray_Group_Jac
```

# Testing select Calf Samples

## Archaea MaTC Jaccard

```{r}
# Metadata 
meta_test <- metadata %>% 
  filter(Animal_Code %in% c("MaTC","MaTD")) 
#   arrange(Fastq_name) %>% 
#   as.data.frame()
# row.names(meta_test) <- meta_test$Fastq_name


Selected_arch <- abund_spec_arch_rare[row.names(abund_spec_arch_rare) %in% meta_test$Fastq_name, ] %>% 
  .[, colSums(. != 0) != 0] # remove species that have 0 total reads counts in the new data set

# Same order? - YES
identical(rownames(meta_test), rownames(Selected_arch))

# Calculate the distance
data_beta_Jac_arch <- beta_get(Selected_arch, method = "jaccard", binary = TRUE)

# Plot
# Jaccard
bray_ordination <- pcoa(data_beta_Jac_arch)
points <- bray_ordination$vectors
points2 <- cbind(points, meta_test)
# Get variances for each dimension
per_variance <- bray_ordination[["values"]][["Relative_eig"]]*100 

# Plot Group
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.2, 
                                  color = Group)) + 
  geom_point(size = 3) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.2 [", round(per_variance[2], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  scale_color_manual(values = c("blue4","orange2")) + 
  theme(title = element_text(size = 18),
        axis.title = element_text(size = 16), 
        axis.text = element_text(size = 14), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 13),
        legend.position = "none",
        plot.title = element_markdown()) + 
  geom_text_repel(
    aes(label = Sample_ID),
    size = 5,
    box.padding = unit(0.4, "lines"),
    point.padding = unit(0.3, "lines"),
    force = 1,   # pushes labels away from points
    min.segment.length = 0 ) # ensures a segment is always drawn)
  
plot_bray_Group_Jac


# Different dimensions
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.3, 
                                  color = Group)) + 
  geom_point(size = 3) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.3 [", round(per_variance[3], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  scale_color_manual(values = c("blue4","orange2")) + 
  theme(title = element_text(size = 18),
        axis.title = element_text(size = 16), 
        axis.text = element_text(size = 14), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 13),
        legend.position = "none",
        plot.title = element_markdown()) + 
  geom_text_repel(
    aes(label = Sample_ID),
    size = 5,
    box.padding = unit(0.4, "lines"),
    point.padding = unit(0.3, "lines"),
    force = 1,   # pushes labels away from points
    min.segment.length = 0 ) # ensures a segment is always drawn)
  
plot_bray_Group_Jac


# Different dimensions
plot_bray_Group_Jac <- ggplot(points2, 
                              aes(x = Axis.1, 
                                  y = Axis.4, 
                                  color = Group)) + 
  geom_point(size = 3) +
  xlab(paste0("PCoA.1 [", round(per_variance[1], 2), "%]")) +
  ylab(paste0("PCoA.4 [", round(per_variance[4], 2), "%]")) +
  labs(color = "Group") + 
  ggtitle("Jaccard") + 
  scale_color_manual(values = c("blue4","orange2")) + 
  theme(title = element_text(size = 18),
        axis.title = element_text(size = 16), 
        axis.text = element_text(size = 14), 
        legend.title = element_text(size = 15), 
        legend.text = element_text(size = 13),
        legend.position = "none",
        plot.title = element_markdown()) + 
  geom_text_repel(
    aes(label = Sample_ID),
    size = 5,
    box.padding = unit(0.4, "lines"),
    point.padding = unit(0.3, "lines"),
    force = 1,   # pushes labels away from points
    min.segment.length = 0 ) # ensures a segment is always drawn)
  
plot_bray_Group_Jac

```
