# Boiler Plate

```{r}
library(ggplot2)
library(vegan) 
#library(car) # Homoscedasticity test
#library(ggpubr) # Density plots
library(hilldiv) # tss()
library(ape) # pcoa
library(lme4) # lmer
library(lmtest) # anova() will give p-value
# library(ggeffects) # ggpredict
#library(readxl)
# library(bestNormalize) # Yeojohnson
# library(Kendall) # MannKendall
library(ggtext)
library(patchwork)
library(lubridate)
```

# Set-up input and calculate alpha diversity measures for each data set

```{r}
# Factor the Group variable so that the box plots are ordered the way I want
metadata_mt_select <- metadata_mt_select %>%
  mutate(Specific_Group = factor(Specific_Group, levels = c("Wild", "Ex-situ"))) 

# Calculate diversity and richness for the three domains in mountain tapirs select
metadata_mt_select$rich_bac <- specnumber(abund_bac_mt_select_rare)
metadata_mt_select$shannon_bac <- diversity(abund_bac_mt_select_rare, index = "shannon")
metadata_mt_select$simpson_bac <- diversity(abund_bac_mt_select_rare, index = "simpson")

metadata_mt_select$rich_arch <- specnumber(abund_arch_mt_select_rare)
metadata_mt_select$shannon_arch <- diversity(abund_arch_mt_select_rare, index = "shannon")
metadata_mt_select$simpson_arch <- diversity(abund_arch_mt_select_rare, index = "simpson")

metadata_mt_select$rich_fun <- specnumber(abund_fun_mt_select_rare)
metadata_mt_select$shannon_fun <- diversity(abund_fun_mt_select_rare, index = "shannon")
metadata_mt_select$simpson_fun <- diversity(abund_fun_mt_select_rare, index = "simpson")

# Set color vector
mt_select_col <- c("Wild" = "#99FF33", "Ex-situ" = "#A6A6A6")

```

# Mountain tapir Wild v. *Ex-situ* alpha diversity assessment (select samples)

## Bacteria

### Set up the variables, run the Wilcox test, and make the plots

```{r}
# Richness

# Run statistical test
# Are the medians different between wild and ex-situ?
wilcox.test(rich_bac ~ Specific_Group, 
            data = metadata_mt_select)

# Create box plot
p1 <- ggplot(data = metadata_mt_select, 
       aes(x = Specific_Group, y = rich_bac, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "Group", values = mt_select_col,
                    breaks = c("Wild", "Ex-situ"),
                    labels = c(
                      "Wild" = "Wild",
                      "Ex-situ" = expression(italic("Ex-situ"))
                    )) +
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title = element_markdown()
  ) + 
  labs(title = "Richness", x = "Group") + 
  ylab("Richness")
p1


# Shannon

# Run statistical test
# Are the medians different between wild and ex-situ?
wilcox.test(shannon_bac ~ Specific_Group, 
            data = metadata_mt_select)
# Create box plot
p2 <- ggplot(data = metadata_mt_select, 
       aes(x = Specific_Group, y = shannon_bac, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "Group", values = mt_select_col,
                    breaks = c("Wild", "Ex-situ"),
                    labels = c(
                      "Wild" = "Wild",
                      "Ex-situ" = expression(italic("Ex-situ"))
                    )) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "bottom",
    plot.title = element_markdown()
  ) + 
  labs(title = "Shannon", x = "Group") + 
  ylab("Shannon")
p2

# Simpson

# Run statistical test
# Are the medians different between wild and ex-situ?
wilcox.test(simpson_bac ~ Specific_Group, 
            data = metadata_mt_select)
# Create box plot
p3 <- ggplot(data = metadata_mt_select, 
       aes(x = Specific_Group, y = simpson_bac, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "Group", values = mt_select_col,
                    breaks = c("Wild", "Ex-situ"),
                    labels = c(
                      "Wild" = "Wild",
                      "Ex-situ" = expression(italic("Ex-situ"))
                    )) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title = element_markdown()
  ) + 
  labs(title = "Simpson", x = "Group") + 
  ylab("Simpson")
p3
```

### Make and save the big plot

```{r}
(p1 | p2 | p3) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 


# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/alpha_bac_species_rare.pdf",
#        width=6,
#        height=3.4)
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/alpha_bac_species_rare.tiff",
#        width=6,
#        height=3.4)
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/alpha_bac_species_rare.png",
#        width=6,
#        height=3.4)
```

## Archaea

### Set up the variables, run the Wilcox test, and make the plots

```{r}
# Richness

# Run statistical test
# Are the medians different between wild and ex-situ?
wilcox.test(rich_arch ~ Specific_Group, 
            data = metadata_mt_select)

# Create box plot
p1 <- ggplot(data = metadata_mt_select, 
       aes(x = Specific_Group, y = rich_arch, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "Group", values = mt_select_col,
                    breaks = c("Wild", "Ex-situ"),
                    labels = c(
                      "Wild" = "Wild",
                      "Ex-situ" = expression(italic("Ex-situ"))
                    )) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title = element_markdown()
  ) + 
  labs(title = "Richness", x = "Group") + 
  ylab("Richness")
p1


# Shannon

# Run statistical test
# Are the medians different between wild and ex-situ?
wilcox.test(shannon_arch ~ Specific_Group, 
            data = metadata_mt_select)
# Create box plot
p2 <- ggplot(data = metadata_mt_select, 
       aes(x = Specific_Group, y = shannon_arch, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "Group", values = mt_select_col,
                    breaks = c("Wild", "Ex-situ"),
                    labels = c(
                      "Wild" = "Wild",
                      "Ex-situ" = expression(italic("Ex-situ"))
                    )) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "bottom",
    plot.title = element_markdown()
  ) + 
  labs(title = "Shannon", x = "Group") + 
  ylab("Shannon")
p2

# Simpson

# Run statistical test
# Are the medians different between wild and ex-situ?
wilcox.test(simpson_arch ~ Specific_Group, 
            data = metadata_mt_select)
# Create box plot
p3 <- ggplot(data = metadata_mt_select, 
       aes(x = Specific_Group, y = simpson_arch, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "Group", values = mt_select_col,
                    breaks = c("Wild", "Ex-situ"),
                    labels = c(
                      "Wild" = "Wild",
                      "Ex-situ" = expression(italic("Ex-situ"))
                    )) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title = element_markdown()
  ) + 
  labs(title = "Simpson", x = "Group") + 
  ylab("Simpson")
p3
```

### Make and save the big plot

```{r}
(p1 | p2 | p3) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 


# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/alpha_arch_species_rare.pdf",
#        width=6,
#        height=3.4)
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/alpha_arch_species_rare.tiff",
#        width=6,
#        height=3.4)
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/alpha_arch_species_rare.png",
#        width=6,
#        height=3.4)
```

## Fungi

### Set up the variables, run the Wilcox test, and make the plots

```{r}
# Richness

# Run statistical test
# Are the medians different between wild and ex-situ?
wilcox.test(rich_fun ~ Specific_Group, 
            data = metadata_mt_select)

# Create box plot
p1 <- ggplot(data = metadata_mt_select, 
       aes(x = Specific_Group, y = rich_fun, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "Group", values = mt_select_col,
                    breaks = c("Wild", "Ex-situ"),
                    labels = c(
                      "Wild" = "Wild",
                      "Ex-situ" = expression(italic("Ex-situ"))
                    )) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title = element_markdown()
  ) + 
  labs(title = "Richness", x = "Group") + 
  ylab("Richness")
p1


# Shannon

# Run statistical test
# Are the medians different between wild and ex-situ?
wilcox.test(shannon_fun ~ Specific_Group, 
            data = metadata_mt_select)
# Create box plot
p2 <- ggplot(data = metadata_mt_select, 
       aes(x = Specific_Group, y = shannon_fun, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "Group", values = mt_select_col,
                    breaks = c("Wild", "Ex-situ"),
                    labels = c(
                      "Wild" = "Wild",
                      "Ex-situ" = expression(italic("Ex-situ"))
                    )) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "bottom",
    plot.title = element_markdown()
  ) + 
  labs(title = "Shannon", x = "Group") + 
  ylab("Shannon")
p2

# Simpson

# Run statistical test
# Are the medians different between wild and ex-situ?
wilcox.test(simpson_fun ~ Specific_Group, 
            data = metadata_mt_select)
# Create box plot
p3 <- ggplot(data = metadata_mt_select, 
       aes(x = Specific_Group, y = simpson_fun, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "Group", values = mt_select_col,
                    breaks = c("Wild", "Ex-situ"),
                    labels = c(
                      "Wild" = "Wild",
                      "Ex-situ" = expression(italic("Ex-situ"))
                    )) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title = element_markdown()
  ) + 
  labs(title = "Simpson", x = "Group") + 
  ylab("Simpson")
p3
```

### Make and save the big plot

```{r}
(p1 | p2 | p3) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 


# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/alpha_fun_species_rare.pdf",
#        width=6,
#        height=3.4)
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/alpha_fun_species_rare.tiff",
#        width=6,
#        height=3.4)
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/alpha_fun_species_rare.png",
#        width=6,
#        height=3.4)
```

# *Ex-situ* Mt. Tapir alpha diversity over time

Plot function for mountain tapirs

```{r}
# Plot function for mountain tapirs
plot_mt <- function (data, variable) {
  ggplot(data = data, 
             aes(x = Date_collected, 
                 y = .data[[variable]], 
                 color = Animal_Code)) + 
  geom_point(aes(shape = Zoo_Code), 
             size = 2.5) +
  geom_line(linewidth = 0.75) + 
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    # scale_x_continuous(breaks = seq(1,9, by  = 2), 
    #                    limits = c(1,9)) +
  scale_color_manual(name = "Animal", 
                     values = c("#619cff","#b79f00",
                                "#7CAE00","yellow3","#F564E3", # yellow3 is #CDCD00
                                "#00b4f0","#de8c00")) +
  scale_shape_manual(values = c(15,19)) +
  guides(color = guide_legend(
    override.aes = list(
      shape = c(19,19,15,15,15,15,15)
    ))) + 
  theme(title = element_text(size = 18),
        plot.title = element_text(hjust = 0.5),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15),
        legend.text = element_text(size = 13), 
        legend.key.width = unit(2, "cm")) 
}
```

### Set up variables and calculate diversity

```{r}
# Calculate diversity and richness for the three domains in mountain tapirs select
metadata_mt_exsitu$rich_bac <- specnumber(abund_bac_mt_exsitu_rare)
metadata_mt_exsitu$shannon_bac <- diversity(abund_bac_mt_exsitu_rare, index = "shannon")
metadata_mt_exsitu$simpson_bac <- diversity(abund_bac_mt_exsitu_rare, index = "simpson")

metadata_mt_exsitu$rich_arch <- specnumber(abund_arch_mt_exsitu_rare)
metadata_mt_exsitu$shannon_arch <- diversity(abund_arch_mt_exsitu_rare, index = "shannon")
metadata_mt_exsitu$simpson_arch <- diversity(abund_arch_mt_exsitu_rare, index = "simpson")

metadata_mt_exsitu$rich_fun <- specnumber(abund_fun_mt_exsitu_rare)
metadata_mt_exsitu$shannon_fun <- diversity(abund_fun_mt_exsitu_rare, index = "shannon")
metadata_mt_exsitu$simpson_fun <- diversity(abund_fun_mt_exsitu_rare, index = "simpson")

# Set color vector

# Fix the date variable so I can actually graph with it
meta_long <- metadata_mt_exsitu %>% 
  mutate(Date_collected = as.Date(.$Date_collected, format = "%m/%d/%Y")) %>% 
  mutate(months_since_start = (time_length(interval(min(Date_collected), Date_collected), unit = "months") + 1))

```

### Make the figures

```{r}

# Richness
# bacteria 
p1 <- plot_mt(data = meta_long, variable = "rich_bac") + 
  ylim(11600,12450) + 
  ylab("Species Richness") + 
  xlab("") + 
  ggtitle("Bacteria") 
p1

# archaea
p2 <- plot_mt(data = meta_long, variable = "rich_arch") + 
  ylim(400,515) +
  ylab("") + 
  xlab("") + 
  ggtitle("Archaea") 
p2

# fungal
p3 <- plot_mt(data = meta_long, variable = "rich_fun") + 
  ylim(115,150) + 
  ylab("") + 
  xlab("") +
  ggtitle("Fungi") 
p3


# Shannon
# bacteria 
p4 <- plot_mt(data = meta_long, variable = "shannon_bac") + 
  ylim(2,8) + 
  ylab("Shannon Diversity") + 
  xlab("") 
p4

# archaea
p5 <- plot_mt(data = meta_long, variable = "shannon_arch") + 
  ylim(2,8) + 
  ylab("") + 
  xlab("") 
p5

# fungal
p6 <- plot_mt(data = meta_long, variable = "shannon_fun") + 
  ylim(2,8) + 
  ylab("") + 
  xlab("") 
p6


# Simpson
# bacteria 
p7 <- plot_mt(data = meta_long, variable = "simpson_bac") + 
  ylim(0.55,1) + 
  ylab("Simpson Diversity") + 
  xlab("") 
p7

# archaea
p8 <- plot_mt(data = meta_long, variable = "simpson_arch") + 
  ylim(0.55,1) + 
  ylab("") + 
  xlab("Date") 
p8

# fungal
p9 <- plot_mt(data = meta_long, variable = "simpson_fun") + 
  ylim(0.55,1) + 
  ylab("") + 
  xlab("") 
p9


# Mega panel
((p1 | p2 | p3) / (p4 | p5 | p6 ) / (p7 | p8 | p9)) + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = "A", title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        legend.position = "right",
        plot.title= element_markdown()) 


# # Save the figure
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/Mountain_alpha_9.tiff",
#        width = 1725,
#        height = 894,
#        units = "px",
#        dpi = 96.14)
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/Mountain_alpha_9.png",
#        width = 1725,
#        height = 894,
#        units = "px",
#        dpi = 96.14)

# The 3 panel figure
((p4 + ggtitle("Bacteria")) | (p5 + ggtitle("Archaea") + xlab("Date")) | (p6 + ggtitle("Fungi")) ) + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = "A", title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        legend.position = "right",
        plot.title= element_markdown(), 
        axis.text.x = element_text(angle = 45, vjust = 0.5),
        panel.grid.minor = element_blank())

# # Save the figure
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/Mountain_alpha_3_shannon.tiff",
#        width = 14,
#        height = 5,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/Mountain_alpha_3_shannon.png",
#        width = 14,
#        height = 5,
#        units = "in")
```

# Set up Dam-Calf data

```{r}
# Calculate diversity and richness for the three domains in dam-calf-sire samples
metadata_dcs$rich_bac <- specnumber(abund_bac_dcs_rare)
metadata_dcs$shannon_bac <- diversity(abund_bac_dcs_rare, index = "shannon")
metadata_dcs$simpson_bac <- diversity(abund_bac_dcs_rare, index = "simpson")

metadata_dcs$rich_arch <- specnumber(abund_arch_dcs_rare)
metadata_dcs$shannon_arch <- diversity(abund_arch_dcs_rare, index = "shannon")
metadata_dcs$simpson_arch <- diversity(abund_arch_dcs_rare, index = "simpson")

metadata_dcs$rich_fun <- specnumber(abund_fun_dcs_rare)
metadata_dcs$shannon_fun <- diversity(abund_fun_dcs_rare, index = "shannon")
metadata_dcs$simpson_fun <- diversity(abund_fun_dcs_rare, index = "simpson")

```

# Malayan tapir calf-dam over time

Plot function for Malayan tapirs.

```{r}
# Plot function
plot_Mal <- function (data, variable) {
  ggplot(data = data, 
             aes(x = months_since_start, 
                 y = .data[[variable]], 
                 color = Animal_Code)) +
  geom_point(size = 4, shape = 3) +
  geom_line(aes(linetype = Group), 
            linewidth = 1.25) + 
  scale_x_continuous(breaks = seq(0,10, by = 1),
                     minor_breaks = NULL) +
  scale_linetype_manual(values = c(2,1), 
                        labels = c("Calf", "Adult"), 
                        name = "Age Group") + 
  scale_color_manual(name = "Animal", 
                     values = c("#00bfc4","#ff64b0", "black")) +
  guides(color = guide_legend(
    override.aes = list(
      shape = c(3,3,3),
      linetype = c(2,1,0)
    ), 
    order = 2),
    linetype = guide_legend(order = 1)) + 
  theme(title = element_text(size = 18),
        plot.title = element_text(hjust = 0.5),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15),
        legend.text = element_text(size = 13), 
        legend.key.width = unit(2, "cm"))
}
```

```{r}
# Fix the date variable so I can actually graph with it and select Malayan Tapirs
meta_long <- metadata_dcs %>% 
  filter(Tapir == "Malayan") %>% 
  mutate(Date_collected = as.Date(.$Date_collected, format = "%m/%d/%Y")) %>% 
  mutate(months_since_start = time_length(interval(min(Date_collected), Date_collected), "months")  + 0.76)

plot_Mal(data = meta_long, variable = "shannon_bac")

# Richness
# Bacteria 
p1 <- plot_Mal(data = meta_long, variable = "rich_bac") + 
  ylab("Species Richness") + 
  xlab("") + 
  ggtitle("Bacteria") #+ 
  #ylim(9000,12000)
p1
# Archaea 
p2 <- plot_Mal(data = meta_long, variable = "rich_arch") + 
  ylab("") + 
  xlab("") + 
  ggtitle("Archaea") #+ 
  #ylim(50,400)
p2
# Fungi
p3 <- plot_Mal(data = meta_long, variable = "rich_fun") + 
  ylab("") + 
  xlab("") + 
  ggtitle("Fungi") + 
  ylim(75,150)
p3


# Shannon
# Bacteria 
p4 <- plot_Mal(data = meta_long, variable = "shannon_bac") + 
  ylab("Shannon Diversity") + 
  xlab("") + 
  ggtitle("") + 
  ylim(2,8)
p4
# Archaea 
p5 <- plot_Mal(data = meta_long, variable = "shannon_arch") + 
  ylab("") + 
  xlab("") + 
  ggtitle("") + 
  ylim(2,8)
p5
# Fungi
p6 <- plot_Mal(data = meta_long, variable = "shannon_fun") + 
  ylab("") + 
  xlab("") + 
  ggtitle("") + 
  ylim(2,8)
p6


# Shannon
# Bacteria 
p7 <- plot_Mal(data = meta_long, variable = "simpson_bac") + 
  ylab("Simpson Diversity") + 
  xlab("") + 
  ggtitle("") + 
  ylim(0.7,1)
p7
# Archaea 
p8 <- plot_Mal(data = meta_long, variable = "simpson_arch") + 
  ylab("") + 
  xlab("Calf age (month of life)") + 
  ggtitle("") + 
  ylim(0.7,1)
p8
# Fungi
p9 <- plot_Mal(data = meta_long, variable = "simpson_fun") + 
  ylab("") + 
  xlab("") + 
  ggtitle("") + 
  ylim(0.7,1)
p9


# Mega panel
((p1 | p2 | p3 ) / (p4 | p5 | p6) / (p7 | p8 | p9)) + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = "A", title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        legend.position = "right",
        plot.title= element_markdown())  


# # Save the figure
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/Malayan_alpha_9.tiff", 
#        width = 1725,
#        height = 894,
#        units = "px",
#        dpi = 96.14)
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/Malayan_alpha_9.png", 
#        width = 1725,
#        height = 894,
#        units = "px",
#        dpi = 96.14)

# The panel we are actually going to use
((p4 + ggtitle("Bacteria")) | (p5 + ggtitle("Archaea") + xlab("Calf age (month of life)")) | (p6 + ggtitle("Fungi")) ) + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = "A", title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        legend.position = "right",
        plot.title= element_markdown()) 

# # Save the figure
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/Malayan_alpha_3_shannon.tiff", 
#        width = 14,
#        height = 5,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/Malayan_alpha_3_shannon.png", 
#        width = 14,
#        height = 5,
#        units = "in")
```

# Lowland tapir calf -dam over time

Plot function for lowland tapirs

```{r}
# Plot function
plot_Low <- function (data, variable) {
  ggplot(data = data, 
             aes(x = months_since_start, 
                 y = .data[[variable]], 
                 color = Animal_Code)) +
  geom_point(size = 4, shape = 17) +
  geom_line(aes(linetype = Group), 
            linewidth = 1.25) + 
  scale_linetype_manual(values = c(2,1), 
                        labels = c("Calf", "Adult"), 
                        name = "Age Group") + 
  scale_color_manual(name = "Animal", 
                     values = c("#c77cff","#00c08b")) +
  guides(color = guide_legend(
    override.aes = list(
      shape = c(17,17),
      linetype = c(2,1)
    ), 
    order = 2), 
    linetype = guide_legend(order = 1)) + 
  theme(title = element_text(size = 18),
        plot.title = element_text(hjust = 0.5),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15),
        legend.text = element_text(size = 13), 
        legend.key.width = unit(2, "cm")) + 
    xlim(0,5)
}

```

```{r}
# Fix the date variable so I can actually graph with it
meta_long <- metadata_dcs %>% 
  filter(Tapir == "Lowland") %>% 
  mutate(Date_collected = as.Date(.$Date_collected, format = "%m/%d/%Y")) %>% 
  mutate(months_since_start = time_length(interval(min(Date_collected), Date_collected), "months") + 0.613)


# Richness
# Bacteria 
p1 <- plot_Low(data = meta_long, variable = "rich_bac") + 
  ylab("Species Richness") + 
  xlab("") + 
  ggtitle("Bacteria") 
p1
# Archaea 
p2 <- plot_Low(data = meta_long, variable = "rich_arch") + 
  ylab("") + 
  xlab("") + 
  ggtitle("Archaea") 
p2
# Fungi 
p3 <- plot_Low(data = meta_long, variable = "rich_fun") + 
  ylab("") + 
  xlab("") + 
  ggtitle("Fungi") 
p3

# Shannon
# Bacteria 
p4 <- plot_Low(data = meta_long, variable = "shannon_bac") + 
  ylab("Shannon Diversity") + 
  xlab("") + 
  ggtitle("") + 
  ylim(0,8) 
p4
# Archaea 
p5 <- plot_Low(data = meta_long, variable = "shannon_arch") + 
  ylab("") + 
  xlab("") + 
  ggtitle("") + 
  ylim(0,8)
p5
# Fungi
p6 <- plot_Low(data = meta_long, variable = "shannon_fun") + 
  ylab("") + 
  xlab("") + 
  ggtitle("") + 
  ylim(0,8)
p6

# Simpson
# Bacteria 
p7 <- plot_Low(data = meta_long, variable = "simpson_bac") + 
  ylab("Simpson Diversity") + 
  xlab("") + 
  ggtitle("") + 
  ylim(0,1)
p7
# Archaea
p8 <- plot_Low(data = meta_long, variable = "simpson_arch") + 
  ylab("") + 
  xlab("Calf age (month of life)") + 
  ggtitle("") + 
  ylim(0,1)
p8
# Fungi
p9 <- plot_Low(data = meta_long, variable = "simpson_fun") + 
  ylab("") + 
  xlab("") + 
  ggtitle("") +
  ylim(0,1)
p9


# Mega panel
((p1 | p2 | p3 ) / (p4 | p5 | p6) / (p7 | p8 | p9)) + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = "A", title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        legend.position = "right",
        plot.title= element_markdown())  


# # Save the plots
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/Lowland_alpha_9.tiff",
#        width = 1725,
#        height = 894,
#        units = "px",
#        dpi = 96.14)
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/Lowland_alpha_9.png",
#        width = 1725,
#        height = 894,
#        units = "px",
#        dpi = 96.14)


# The panel we are actually going to use
((p4 + ggtitle("Bacteria")) | (p5 + ggtitle("Archaea") + xlab("Calf age (month of life)")) | (p6 + ggtitle("Fungi")) ) + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = "A", title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        legend.position = "right",
        plot.title= element_markdown()) 

# # Save the figure
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/Lowland_alpha_3_shannon.tiff",
#        width = 14,
#        height = 5,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/All MT samples Experimental separation rarefaction/Lowland_alpha_3_shannon.png",
#        width = 14,
#        height = 5,
#        units = "in")
```
