# Boiler Plate

```{r}
library(ggplot2)
library(vegan) 
#library(car) # Homoscedasticity test
#library(ggpubr) # Density plots
library(hilldiv) # tss()
library(ape) # pcoa
library(lme4) # lmer
library(lmtest) # anova() will give p-value
# library(ggeffects) # ggpredict
#library(readxl)
# library(bestNormalize) # Yeojohnson
# library(Kendall) # MannKendall
library(ggtext)
library(patchwork)
library(lubridate)
```

# Functions

```{r}
`%notin%` <- function (x, y) {!(x %in% y)}
```

### Species

```{r}
# Returns a data frame with only specific samples based on metadata values
select_samples <- function(data, metadata, group = "", values = c()) {
  hold <- metadata$Fastq_name[metadata[[group]] %in% values]
  data_subset <- data %>% 
    t() %>% 
    as.data.frame() %>%
    select(any_of(hold)) %>% 
    t() %>% 
    as.data.frame()
  
  return(data_subset)
}
```

# Set-up input and calculate Shannon diversity

Select input data frame and calculate Shannon diversity (store in metadata variable).

```{r}
# Grab the proper data frame to do everything beyond this point
# This way if I chnage the data frame I am using I only need to rewrite 
# 1 line of code
# data_Div <- kraken_130_species
# 
# # Calculate shannon diversity
# metadata$shannon <- NA
# a_div <- diversity(data_Div, index = "shannon")
# 
# # Metadata and abundance data rows are NOT in the same order, thus I need to match each diversity value with
# # the correct corresponding row in metadata. Otherwise is asigned incorrectly.
# for (i in names(a_div)) {
#   metadata[metadata$Fastq_name == as.character(i),"shannon"] <- a_div[as.character(i)]
# }
# rm(a_div)


# Calculate diversity and richness for the three domains
metadata$shannon_bac <- diversity(abund_spec_bac, index = "shannon")
metadata$simpson_bac <- diversity(abund_spec_bac, index = "simpson")
metadata$rich_bac <- specnumber(abund_spec_bac)

metadata$shannon_arch <- diversity(abund_spec_arch, index = "shannon")
metadata$simpson_arch <- diversity(abund_spec_arch, index = "simpson")
metadata$rich_arch <- specnumber(abund_spec_arch)

metadata$shannon_fun <- diversity(abund_spec_fun, index = "shannon")
metadata$simpson_fun <- diversity(abund_spec_fun, index = "simpson")
metadata$rich_fun <- specnumber(abund_spec_fun)

```

Rarefaction for each of archaea, bacteria, fungi

```{r}
# # Find the minimum library size (the sample with the fewest reads)
# raremax <- min(rowSums(abund_spec_bac))
# # Rarify the data (this takes a minute)
# set.seed(1025)
# abund_spec_bac_rare <- rrarefy(abund_spec_bac, raremax) %>% 
#   as.data.frame()
# 
# # Find the minimum library size (the sample with the fewest reads)
# raremax <- min(rowSums(abund_spec_arch))
# # Rarify the data (this takes a minute)
# set.seed(1025)
# abund_spec_arch_rare <- rrarefy(abund_spec_arch, raremax) %>% 
#   as.data.frame
# 
# # Find the minimum library size (the sample with the fewest reads)
# raremax <- min(rowSums(abund_spec_fun))
# # Rarify the data (this takes a minute)
# set.seed(1025)
# abund_spec_fun_rare <- rrarefy(abund_spec_fun, raremax) %>% 
#   as.data.frame()
# 
# # Check this didn't ruin the row names and order
# identical(rownames(abund_spec_bac_rare), rownames(metadata))
# identical(rownames(abund_spec_fun_rare), rownames(metadata))
# identical(rownames(abund_spec_arch_rare), rownames(metadata))

# Once again double check that the smaples are in the same order, otherwise everything will be mismatched ALL TRUE
identical(rownames(metadata), rownames(abund_spec_bac_rare))
identical(rownames(metadata), rownames(abund_spec_arch_rare))
identical(rownames(metadata), rownames(abund_spec_fun_rare))
identical(rownames(metadata), rownames(abund_spec_virus_rare))

# identical(rownames(metadata), rownames(abund_spec_bac))
# identical(rownames(metadata), rownames(abund_spec_arch))
# identical(rownames(metadata), rownames(abund_spec_fun))

#metadata <- metadata %>% filter(Tapir == "Mountain") # For testing the new rarified abundances

# Calculate diversity on rareified data
metadata$shannon_bac <- diversity(abund_spec_bac_rare, index = "shannon")
metadata$simpson_bac <- diversity(abund_spec_bac_rare, index = "simpson")
metadata$rich_bac <- specnumber(abund_spec_bac_rare)
# metadata$bac_OG <- specnumber(abund_spec_bac)

metadata$shannon_arch <- diversity(abund_spec_arch_rare, index = "shannon")
metadata$simpson_arch <- diversity(abund_spec_arch_rare, index = "simpson")
metadata$rich_arch <- specnumber(abund_spec_arch_rare)
# metadata$arch_OG <- specnumber(abund_spec_arch)

metadata$shannon_fun <- diversity(abund_spec_fun_rare, index = "shannon")
metadata$simpson_fun <- diversity(abund_spec_fun_rare, index = "simpson")
metadata$rich_fun <- specnumber(abund_spec_fun_rare)
# metadata$fun_OG <- specnumber(abund_spec_fun)

metadata$shannon_virus <- diversity(abund_spec_virus_rare, index = "shannon")
metadata$simpson_virus <- diversity(abund_spec_virus_rare, index = "simpson")
metadata$rich_virus <- specnumber(abund_spec_virus_rare)
# metadata$fun_OG <- specnumber(abund_spec_fun)
```

# Captive v. Wild

## Use all mountain tapir samples and account for replicates

```{r}
# Grab the data
meta_test <- metadata %>% 
  filter(Specific_Group %in% c("Wild","Ex-situ")) %>% 
  mutate(Specific_Group = factor(.$Specific_Group, levels = c("Wild", "Ex-situ")))

## Bacteria
print("Bacteria ================================================================================")
# Make the Sattertwaite method
lmm_model_bac_rich_rare <- lmer(rich_bac ~ Specific_Group + (1|Animal_Code),
                          data = meta_test)
# Get the p-value using the anova() method aliReza showed me
lmm_model_absent_rich_rare <- lmer(rich_bac ~ 1 + (1|Animal_Code),
                         data = meta_test)
print("p-value by ANOVA comparison of two LMM models")
anova(lmm_model_absent_rich_rare, lmm_model_bac_rich_rare) # To get AIC

# Make the Sattertwaite method
lmm_model_bac_shan_rare <- lmer(shannon_bac ~ Specific_Group + (1|Animal_Code),
                          data = meta_test)
# Get the p-value using the anova() method aliReza showed me
lmm_model_absent_shan_rare <- lmer(shannon_bac ~ 1 + (1|Animal_Code),
                         data = meta_test)
print("p-value by ANOVA comparison of two LMM models")
anova(lmm_model_absent_shan_rare, lmm_model_bac_shan_rare) # To get AIC

# Make the Sattertwaite method
lmm_model_bac_simp_rare <- lmer(simpson_bac ~ Specific_Group + (1|Animal_Code),
                          data = meta_test)
# Get the p-value using the anova() method aliReza showed me
lmm_model_absent_simp_rare <- lmer(simpson_bac ~ 1 + (1|Animal_Code),
                         data = meta_test)
print("p-value by ANOVA comparison of two LMM models")
anova(lmm_model_absent_simp_rare, lmm_model_bac_simp_rare) # To get AIC

print("Archaea ================================================================================")
## Archaea
# Make the Sattertwaite method
lmm_model_arch_rich_rare <- lmer(rich_arch ~ Specific_Group + (1|Animal_Code),
                          data = meta_test)
# Get the p-value using the anova() method aliReza showed me
lmm_model_absent_rich_rare <- lmer(rich_arch ~ 1 + (1|Animal_Code),
                         data = meta_test)
print("p-value by ANOVA comparison of two LMM models")
anova(lmm_model_absent_rich_rare, lmm_model_arch_rich_rare) # To get AIC

# Make the Sattertwaite method
lmm_model_arch_shan_rare <- lmer(shannon_arch ~ Specific_Group + (1|Animal_Code),
                          data = meta_test)
# Get the p-value using the anova() method aliReza showed me
lmm_model_absent_shan_rare <- lmer(shannon_arch ~ 1 + (1|Animal_Code),
                         data = meta_test)
print("p-value by ANOVA comparison of two LMM models")
anova(lmm_model_absent_shan_rare, lmm_model_arch_shan_rare) # To get AIC

# Make the Sattertwaite method
lmm_model_arch_simp_rare <- lmer(simpson_arch ~ Specific_Group + (1|Animal_Code),
                          data = meta_test)
# Get the p-value using the anova() method aliReza showed me
lmm_model_absent_simp_rare <- lmer(simpson_arch ~ 1 + (1|Animal_Code),
                         data = meta_test)
print("p-value by ANOVA comparison of two LMM models")
anova(lmm_model_absent_simp_rare, lmm_model_arch_simp_rare) # To get AIC

print("Fungi ================================================================================")
## Fungi
# Make the Sattertwaite method
lmm_model_fun_rich_rare <- lmer(rich_fun ~ Specific_Group + (1|Animal_Code),
                          data = meta_test)
# Get the p-value using the anova() method aliReza showed me
lmm_model_absent_shan_rare <- lmer(rich_fun ~ 1 + (1|Animal_Code),
                         data = meta_test)
print("p-value by ANOVA comparison of two LMM models")
anova(lmm_model_absent_rich_rare, lmm_model_fun_rich_rare) # To get AIC

# Make the Sattertwaite method
lmm_model_fun_shan_rare <- lmer(shannon_fun ~ Specific_Group + (1|Animal_Code),
                          data = meta_test)
# Get the p-value using the anova() method aliReza showed me
lmm_model_absent_shan_rare <- lmer(shannon_fun ~ 1 + (1|Animal_Code),
                         data = meta_test)
print("p-value by ANOVA comparison of two LMM models")
anova(lmm_model_absent_shan_rare, lmm_model_fun_shan_rare) # To get AIC

# Make the Sattertwaite method
lmm_model_fun_simp_rare <- lmer(simpson_fun ~ Specific_Group + (1|Animal_Code),
                          data = meta_test)
# Get the p-value using the anova() method aliReza showed me
lmm_model_absent_simp_rare <- lmer(simpson_fun ~ 1 + (1|Animal_Code),
                         data = meta_test)
print("p-value by ANOVA comparison of two LMM models")
anova(lmm_model_absent_simp_rare, lmm_model_fun_simp_rare) # To get AIC
```

## Set up specific input data and Test statistical assumptions (OG analysis)

```{r}
meta_test <- metadata %>% 
  filter(Specific_Group %in% c("Wild","Ex-situ")) %>% 
  group_by(Animal_Code) %>% 
  arrange(Date_collected, .by_group = TRUE) %>% 
  slice(1) %>% 
  ungroup() %>% 
  mutate(Specific_Group = factor(.$Specific_Group, levels = c("Wild", "Ex-situ")))

# # Test of normality via Shapiro-Wilks test FAILED 
# shapiro.test(meta_test$shannon) # p-value = 0.0225
# 
# # Visualize normality
# qqnorm(meta_test$shannon) 
# qqline(meta_test$shannon)
# text(qqnorm(meta_test$shannon, plot.it = FALSE), 
#      labels = meta_test$Fastq_name, 
#      pos = 4, cex = 0.7)
# ggdensity(meta_test$shannon)
# 
# # Test of Homoskedasticity via levene test PASS 
# with(meta_test, leveneTest(shannon, Specific_Group)) # p = 0.9664
```

check the sampling depth now that I have rarefied. Compare to un-rarefied richness.

```{r}
ggplot(data = meta_test, aes(x = `QC read count (Millions)`, y = rich_bac, color = Group)) + 
  geom_point()
ggplot(data = meta_test, aes(x = `QC read count (Millions)`, y = bac_OG, color = Group)) + 
  geom_point()

ggplot(data = meta_test, aes(x = `QC read count (Millions)`, y = rich_fun, color = Group)) + 
  geom_point()
ggplot(data = meta_test, aes(x = `QC read count (Millions)`, y = fun_OG, color = Group)) + 
  geom_point()

ggplot(data = meta_test, aes(x = `QC read count (Millions)`, y = rich_arch, color = Group)) + 
  geom_point()
ggplot(data = meta_test, aes(x = `QC read count (Millions)`, y = arch_OG, color = Group)) + 
  geom_point()
```

## Run Wilcox test and create box plot

bacteria

```{r}
# # Testing normality of all the values
# shapiro.test(meta_test$shannon_bac) # PASS
# shapiro.test(meta_test$simpson_bac) # FAIL
# shapiro.test(meta_test$rich_bac) # PASS
# 
# # Testing homoskedasticity of all the values
# library(car)
# with(meta_test, leveneTest(shannon_bac, Specific_Group)) # PASS
# with(meta_test, leveneTest(simpson_bac, Specific_Group)) # PASS 
# with(meta_test, leveneTest(rich_bac, Specific_Group)) # PASS
# 
# # T-test for those values that passed both tests
# t.test(shannon_bac ~ Specific_Group, data = meta_test) # p-value = 0.6614
# t.test(rich_bac ~ Specific_Group, data = meta_test) # p-value = 0.0266

# Wilcox test (the non-parametric version of the independent t test)
wilcox.test(shannon_bac ~ Specific_Group, data = meta_test) # p-value = 0.9452
wilcox.test(simpson_bac ~ Specific_Group, data = meta_test) # p-value = 1
wilcox.test(rich_bac ~ Specific_Group, data = meta_test) # p-value = 0.05128

# Create box plot
p1 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = shannon_bac, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "Group", values = c("#99FF33", "#A6A6A6"),
                    breaks = c("Wild", "Ex-situ"),
                    labels = c(
                      "Wild" = "Wild",
                      "Ex-situ" = expression(italic("Ex-situ"))
                    )) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "bottom",
    plot.title = element_markdown()
  ) + 
  labs(title = "Shannon", x = "Group") + 
  ylab("Shannon")
p1

# Create box plot
p3 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = simpson_bac, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  theme_bw() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  scale_fill_manual(name = "Group", values = c("#99FF33", "#A6A6A6"),
                    breaks = c("Wild", "Ex-situ")) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title= element_markdown()
  ) + 
  labs(title = "Simpson", x = "Group") + 
  ylab("Simpson")
p3

# Create box plot
p2 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = rich_bac, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  theme_bw() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  scale_fill_manual(name = "Group", values = c("#99FF33", "#A6A6A6"),
                    breaks = c("Wild", "Ex-situ")) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title= element_markdown()
  ) + 
  labs(title = "Richness", x = "Group") + 
  ylab("Richness")
p2

(p2 | p1 | p3) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 


# # (p-value = 0.7308) old for allDiv
# ggsave(file = "../Images/Manuscript (Frontiers)/alpha_bac_species_rare.pdf", 
#        width=6, 
#        height=3.4)
# ggsave(file = "../Images/Manuscript (Frontiers)/alpha_bac_species_rare.tiff", 
#        width=6, 
#        height=3.4)
# ggsave(file = "../Images/Manuscript (Frontiers)/alpha_bac_species_rare.png", 
#        width=6, 
#        height=3.4)
```

bacteria for presentation

```{r}
p1 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = shannon_bac, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "", values = c("#99FF33", "#A6A6A6"),
                    breaks = c("Wild", "Ex-situ")) + 
  theme(
    title = element_text(size = 14),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    # axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title= element_markdown()
  ) + 
  labs(title = "p = 0.94", x = "Group") + 
  ylab("Shannon")
p1

ggsave( file = "../Images/Presentation/Alpha_Bac_Shannon.png", 
        units = "in",
        height = 2.97,
        width = 2.3)
```

Archaea

```{r}
# # Testing normality of all the values
# shapiro.test(meta_test$shannon_arch) # PASS
# shapiro.test(meta_test$simpson_arch) # FAIL
# shapiro.test(meta_test$rich_arch) # PASS
# 
# # Testing homoskedasticity of all the values
# library(car)
# with(meta_test, leveneTest(shannon_arch, Specific_Group)) # PASS
# with(meta_test, leveneTest(simpson_arch, Specific_Group)) # PASS 
# with(meta_test, leveneTest(rich_arch, Specific_Group)) # PASS
# 
# # T-test for those values that passed both tests
# t.test(shannon_arch ~ Specific_Group, data = meta_test) # p-value = 0.05253
# t.test(rich_arch ~ Specific_Group, data = meta_test) # p-value = 0.0002824
# 

# Wilcox test (the non-parametric version of the independent t test)
wilcox.test(shannon_arch ~ Specific_Group, data = meta_test) # p-value = 0.05128
wilcox.test(simpson_arch ~ Specific_Group, data = meta_test) # p-value = 0.02214
wilcox.test(rich_arch ~ Specific_Group, data = meta_test) # p-value = 0.001166

# Create box plot
p1 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = shannon_arch, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  theme_bw() + 
  geom_jitter(width = 0.2, size = 1, color = "black") + 
  scale_fill_manual(name = "Group", values = c("#99FF33", "#A6A6A6"),
                    breaks = c("Wild", "Ex-situ"),
                    labels = c(
                      "Wild" = "Wild",
                      "Ex-situ" = expression(italic("Ex-situ"))
                    )) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "bottom",
    plot.title= element_markdown()
  ) + 
  labs(title = "Shannon", x = "Group") + 
  ylab("Shannon")
p1

# Create box plot
p3 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = simpson_arch, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  theme_bw() + 
  geom_jitter(width = 0.2, size = 1, color = "black") + 
  scale_fill_manual(name = "Group", values = c("#99FF33", "#A6A6A6"),
                    breaks = c("Wild", "Ex-situ")) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title= element_markdown()
  ) + 
  labs(title = "Simpson", x = "Group") + 
  ylab("Simpson")
p3

# Create box plot
p2 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = rich_arch, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  theme_bw() + 
  geom_jitter(width = 0.2, size = 1, color = "black") + 
  scale_fill_manual(name = "Group", values = c("#99FF33", "#A6A6A6"),
                    breaks = c("Wild", "Ex-situ")) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title= element_markdown()
  ) + 
  labs(title = "Richness", x = "Group") + 
  ylab("Richness")
p2

(p2 | p1 | p3) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 

# Save the images
# ggsave(file = "../Images/Manuscript (Frontiers)/alpha_arch_species_rare.pdf", 
#        width=6, 
#        height=3.4)
# ggsave(file = "../Images/Manuscript (Frontiers)/alpha_arch_species_rare.tiff", 
#        width=6, 
#        height=3.4)
# ggsave(file = "../Images/Manuscript (Frontiers)/alpha_arch_species_rare.png", 
#        width=6, 
#        height=3.4)
```

Archaea for presentation

```{r}
p1 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = shannon_arch, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "", values = c("#99FF33", "#A6A6A6"),
                    breaks = c("Wild", "Ex-situ")) + 
  theme(
    title = element_text(size = 14),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    # axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title= element_markdown()
  ) + 
  labs(title = "p = 0.051", x = "Group") + 
  ylab("Shannon")
p1

ggsave( file = "../Images/Presentation/Alpha_Arch_Shannon.png", 
        units = "in",
        height = 2.97,
        width = 2.3)
```

Fungi

```{r}
# # Testing normality of all the values
# shapiro.test(meta_test$shannon_fun) # FAIL
# shapiro.test(meta_test$simpson_fun) # FAIL
# shapiro.test(meta_test$rich_fun) # PASS
# 
# # Testing homoskedasticity of all the values
# library(car)
# with(meta_test, leveneTest(shannon_fun, Specific_Group)) # FAIL
# with(meta_test, leveneTest(simpson_fun, Specific_Group)) # FAIL 
# with(meta_test, leveneTest(rich_fun, Specific_Group)) # PASS
# 
# # T-test for those values that passed both tests
# t.test(rich_fun ~ Specific_Group, data = meta_test) # p-value = 0.8134

# Wilcox test (the non-parametric version of the independent t test)
wilcox.test(shannon_fun ~ Specific_Group, data = meta_test) # p-value = 0.2949
wilcox.test(simpson_fun ~ Specific_Group, data = meta_test) # p-value = 0.4452
wilcox.test(rich_fun ~ Specific_Group, data = meta_test) # p-value = 0.7059

# Create box plot
p1 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = shannon_fun, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  theme_bw() + 
  geom_jitter(width = 0.2, size = 1, color = "black") + 
  scale_fill_manual(name = "Group", 
                    values = c("#99FF33", "#A6A6A6"),
                    breaks = c("Wild", "Ex-situ"),
                    labels = c(
                      "Wild" = "Wild",
                      "Ex-situ" = expression(italic("Ex-situ"))
                    )) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "bottom",
    plot.title= element_markdown()
  ) + 
  labs(title = "Shannon", x = "Group") + 
  ylab("Shannon")
p1

# Create box plot
p3 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = simpson_fun, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  theme_bw() + 
  geom_jitter(width = 0.2, size = 1, color = "black") + 
  scale_fill_manual(name = "Group", values = c("#99FF33", "#A6A6A6"),
                    breaks = c("Wild", "Ex-situ")) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title= element_markdown()
  ) + 
  labs(title = "Simpson", x = "Group") + 
  ylab("Simpson")
p3

# Create box plot
p2 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = rich_fun, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  theme_bw() + 
  geom_jitter(width = 0.2, size = 1, color = "black") + 
  scale_fill_manual(name = "Group", values = c("#99FF33", "#A6A6A6"),
                    breaks = c("Wild", "Ex-situ")) + 
  theme(
    title = element_text(size = 16),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title= element_markdown()
  ) + 
  labs(title = "Richness", x = "Group") + 
  ylab("Richness")
p2

(p2 | p1 | p3) + plot_annotation(tag_levels = "A", 
                             title = "") & 
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 

# Save the images
# ggsave(file = "../Images/Manuscript (Frontiers)/alpha_fun_species_rare.pdf", 
#        width=7.5, 
#        height=4.25)
# ggsave(file = "../Images/Manuscript (Frontiers)/alpha_fun_species_rare.tiff", 
#        width=7.5, 
#        height=4.25)
# ggsave(file = "../Images/Manuscript (Frontiers)/alpha_fun_species_rare.png", 
#        width=7.5, 
#        height=4.25)
```

Fungi for presentation

```{r}
p1 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = shannon_fun, 
           fill = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_fill_manual(name = "", values = c("#99FF33", "#A6A6A6"),
                    breaks = c("Wild", "Ex-situ")) + 
  theme(
    title = element_text(size = 14),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    # axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title= element_markdown()
  ) + 
  labs(title = "p = 0.051", x = "Group") + 
  ylab("Shannon")
p1

ggsave( file = "../Images/Presentation/Alpha_fun_Shannon.png", 
        units = "in",
        height = 2.97,
        width = 2.3)
```

Virus

```{r}
# # Testing normality of all the values
# shapiro.test(meta_test$shannon_bac) # PASS
# shapiro.test(meta_test$simpson_bac) # FAIL
# shapiro.test(meta_test$rich_bac) # PASS
# 
# # Testing homoskedasticity of all the values
# library(car)
# with(meta_test, leveneTest(shannon_bac, Specific_Group)) # PASS
# with(meta_test, leveneTest(simpson_bac, Specific_Group)) # PASS 
# with(meta_test, leveneTest(rich_bac, Specific_Group)) # PASS
# 
# # T-test for those values that passed both tests
# t.test(shannon_bac ~ Specific_Group, data = meta_test) # p-value = 0.6614
# t.test(rich_bac ~ Specific_Group, data = meta_test) # p-value = 0.0266

# Wilcox test (the non-parametric version of the independent t test)
wilcox.test(shannon_virus ~ Specific_Group, data = meta_test) # 
wilcox.test(simpson_virus ~ Specific_Group, data = meta_test) # 
wilcox.test(rich_virus ~ Specific_Group, data = meta_test) # 

# Create box plot
p1 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = shannon_virus, 
           color = Specific_Group)) + 
  geom_boxplot() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  theme_bw() + 
  scale_color_manual(name = "Group", values = c("orange1", "blue4")) + 
  theme(
    title = element_text(size = 18),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "bottom",
    plot.title= element_markdown()
  ) + 
  labs(title = "Shannon", x = "Group") + 
  ylab("Shannon")
p1

# Create box plot
p3 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = simpson_virus, 
           color = Specific_Group)) + 
  geom_boxplot() + 
  theme_bw() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  scale_color_manual(name = "Group", values = c("orange1", "blue4")) + 
  theme(
    title = element_text(size = 18),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title= element_markdown()
  ) + 
  labs(title = "Simpson", x = "Group") + 
  ylab("Simpson")
p3

# Create box plot
p2 <- ggplot(data = meta_test, 
       aes(x = Specific_Group, y = rich_virus, 
           color = Specific_Group)) + 
  geom_boxplot() + 
  theme_bw() + 
  geom_jitter(width = 0.2, height = 0, size = 1, color = "black") + 
  scale_color_manual(name = "Group", values = c("orange1", "blue4")) + 
  theme(
    title = element_text(size = 18),
    axis.title = element_text(size = 16), 
    axis.title.x = element_blank(),
    axis.text = element_text(size = 14),
    axis.text.x = element_blank(),
    legend.title = element_text(size = 15), 
    legend.text = element_text(size = 13), 
    legend.position = "none",
    plot.title= element_markdown()
  ) + 
  labs(title = "Richness", x = "Group") + 
  ylab("Richness")
p2

(p2 | p1 | p3) + plot_annotation(tag_levels = "a", 
                             title = "Wild v. *ex-situ* mt. tapirs (Virus)") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        plot.title= element_markdown()) 
```

# Ex-situ Mt. Tapir alpha diversity over time

The original code for this is located in Specialty_Figures.rmd

```{r}
# Select the data for the ex-situ mountain tapirs
meta_long <- metadata %>% 
  filter(Specific_Group == "Ex-situ") %>% 
  mutate(Date_collected = as.Date(.$Date_collected, format = "%m/%d/%Y")) %>% 
  mutate(months_since_start = (time_length(interval(min(Date_collected), Date_collected), unit = "months") + 1))

# Plot function for mountain tapirs
plot_mt <- function (data, variable) {
  ggplot(data = data, 
             aes(x = Date_collected, 
                 y = .data[[variable]], 
                 color = Animal_Code)) + 
  geom_point(aes(shape = Zoo_Code), 
             size = 2.5) +
  geom_line(linewidth = 0.75) + 
  scale_x_date(date_breaks = "1 month", date_labels = "%b") +
    # scale_x_continuous(breaks = seq(1,9, by  = 2), 
    #                    limits = c(1,9)) +
  scale_color_manual(name = "Animal", 
                     values = c("#619cff","#b79f00",
                                "#7CAE00","yellow3","#F564E3",
                                "#00b4f0","#de8c00")) +
  scale_shape_manual(values = c(15,19)) +
  guides(color = guide_legend(
    override.aes = list(
      shape = c(19,19,15,15,15,15,15)
    ))) + 
  theme(title = element_text(size = 18),
        plot.title = element_text(hjust = 0.5),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15),
        legend.text = element_text(size = 13), 
        legend.key.width = unit(2, "cm")) 
}


# Richness
# bacteria 
p1 <- plot_mt(data = meta_long, variable = "rich_bac") + 
  ylim(11125,12125) + 
  ylab("Species Richness") + 
  xlab("") + 
  ggtitle("Bacteria") 
p1

# archaea
p2 <- plot_mt(data = meta_long, variable = "rich_arch") + 
  ylim(50,400) +
  ylab("") + 
  xlab("") + 
  ggtitle("Archaea") 
p2

# fungal
p3 <- plot_mt(data = meta_long, variable = "rich_fun") + 
  ylim(115,150) + 
  ylab("") + 
  xlab("") +
  ggtitle("Fungi") 
p3


# Shannon
# bacteria 
p4 <- plot_mt(data = meta_long, variable = "shannon_bac") + 
  ylim(2,8) + 
  ylab("Shannon Diversity") + 
  xlab("") 
p4

# archaea
p5 <- plot_mt(data = meta_long, variable = "shannon_arch") + 
  ylim(2,8) + 
  ylab("") + 
  xlab("") 
p5

# fungal
p6 <- plot_mt(data = meta_long, variable = "shannon_fun") + 
  ylim(2,8) + 
  ylab("") + 
  xlab("") 
p6


# Simpson
# bacteria 
p7 <- plot_mt(data = meta_long, variable = "simpson_bac") + 
  ylim(0.55,1) + 
  ylab("Simpson Diversity") + 
  xlab("") 
p7

# archaea
p8 <- plot_mt(data = meta_long, variable = "simpson_arch") + 
  ylim(0.55,1) + 
  ylab("") + 
  xlab("Date") 
p8

# fungal
p9 <- plot_mt(data = meta_long, variable = "simpson_fun") + 
  ylim(0.55,1) + 
  ylab("") + 
  xlab("") 
p9


# Mega panel
((p1 | p2 | p3) / (p4 | p5 | p6 ) / (p7 | p8 | p9)) + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = "A", title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        legend.position = "right",
        plot.title= element_markdown()) 


# # Save the figure
# ggsave(file = "../Images/Manuscript (Frontiers)/Mountain_alpha_9.tiff", 
#        width = 1725,
#        height = 894,
#        units = "px",
#        dpi = 96.14)
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/Mountain_alpha_9.png", 
#        width = 1725,
#        height = 894,
#        units = "px",
#        dpi = 96.14)

# The panel we are actually going to use
((p4 + ggtitle("Bacteria")) | (p5 + ggtitle("Archaea") + xlab("Date")) | (p6 + ggtitle("Fungi")) ) + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = "A", title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        legend.position = "right",
        plot.title= element_markdown(), 
        axis.text.x = element_text(angle = 45, vjust = 0.5),
        panel.grid.minor = element_blank())

# Save the figure
# ggsave(file = "../Images/Manuscript (Frontiers)/Mountain_alpha_3_shannon.tiff", 
#        width = 14,
#        height = 5,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/Mountain_alpha_3_shannon.png", 
#        width = 14,
#        height = 5,
#        units = "in")
```

# Malayan tapir calf-dam over time

```{r}
# Select the data for the ex-situ mountain tapirs
meta_long <- metadata %>% 
  filter(Tapir == "Malayan") %>% 
  mutate(Date_collected = as.Date(.$Date_collected, format = "%m/%d/%Y")) %>% 
  mutate(months_since_start = time_length(interval(min(Date_collected), Date_collected), "months")  + 0.76)

# Plot function
plot_Mal <- function (data, variable) {
  ggplot(data = data, 
             aes(x = months_since_start, 
                 y = .data[[variable]], 
                 color = Animal_Code)) +
  geom_point(size = 4, shape = 3) +
  geom_line(aes(linetype = Group), 
            linewidth = 1.25) + 
  scale_x_continuous(breaks = seq(0,10, by = 1),
                     minor_breaks = NULL) +
  scale_linetype_manual(values = c(2,1), 
                        labels = c("Calf", "Adult"), 
                        name = "Age Group") + 
  scale_color_manual(name = "Animal", 
                     values = c("#00bfc4","#ff64b0", "black")) +
  guides(color = guide_legend(
    override.aes = list(
      shape = c(3,3,3),
      linetype = c(2,1,0)
    ), 
    order = 2),
    linetype = guide_legend(order = 1)) + 
  theme(title = element_text(size = 18),
        plot.title = element_text(hjust = 0.5),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15),
        legend.text = element_text(size = 13), 
        legend.key.width = unit(2, "cm"))
}

plot_Mal(data = meta_long, variable = "shannon_bac")

# Richness
# Bacteria 
p1 <- plot_Mal(data = meta_long, variable = "rich_bac") + 
  ylab("Species Richness") + 
  xlab("") + 
  ggtitle("Bacteria") + 
  ylim(9000,12000)
p1
# Archaea 
p2 <- plot_Mal(data = meta_long, variable = "rich_arch") + 
  ylab("") + 
  xlab("") + 
  ggtitle("Archaea") + 
  ylim(50,400)
p2
# Fungi
p3 <- plot_Mal(data = meta_long, variable = "rich_fun") + 
  ylab("") + 
  xlab("") + 
  ggtitle("Fungi") + 
  ylim(75,150)
p3


# Shannon
# Bacteria 
p4 <- plot_Mal(data = meta_long, variable = "shannon_bac") + 
  ylab("Shannon Diversity") + 
  xlab("") + 
  ggtitle("") + 
  ylim(2,8)
p4
# Archaea 
p5 <- plot_Mal(data = meta_long, variable = "shannon_arch") + 
  ylab("") + 
  xlab("") + 
  ggtitle("") + 
  ylim(2,8)
p5
# Fungi
p6 <- plot_Mal(data = meta_long, variable = "shannon_fun") + 
  ylab("") + 
  xlab("") + 
  ggtitle("") + 
  ylim(2,8)
p6


# Shannon
# Bacteria 
p7 <- plot_Mal(data = meta_long, variable = "simpson_bac") + 
  ylab("Simpson Diversity") + 
  xlab("") + 
  ggtitle("") + 
  ylim(0.7,1)
p7
# Archaea 
p8 <- plot_Mal(data = meta_long, variable = "simpson_arch") + 
  ylab("") + 
  xlab("Calf age (month of life)") + 
  ggtitle("") + 
  ylim(0.7,1)
p8
# Fungi
p9 <- plot_Mal(data = meta_long, variable = "simpson_fun") + 
  ylab("") + 
  xlab("") + 
  ggtitle("") + 
  ylim(0.7,1)
p9


# Mega panel
((p1 | p2 | p3 ) / (p4 | p5 | p6) / (p7 | p8 | p9)) + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = "A", title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        legend.position = "right",
        plot.title= element_markdown())  


# Save the figure
ggsave(file = "../Images/Manuscript (Frontiers)/Malayan_alpha_9.tiff", 
       width = 1725,
       height = 894,
       units = "px",
       dpi = 96.14)

ggsave(file = "../Images/Manuscript (Frontiers)/Malayan_alpha_9.png", 
       width = 1725,
       height = 894,
       units = "px",
       dpi = 96.14)

# The panel we are actually going to use
((p4 + ggtitle("Bacteria")) | (p5 + ggtitle("Archaea") + xlab("Calf age (month of life)")) | (p6 + ggtitle("Fungi")) ) + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = "A", title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        legend.position = "right",
        plot.title= element_markdown()) 

# Save the figure
ggsave(file = "../Images/Manuscript (Frontiers)/Malayan_alpha_3_shannon.tiff", 
       width = 14,
       height = 5,
       units = "in")

ggsave(file = "../Images/Manuscript (Frontiers)/Malayan_alpha_3_shannon.png", 
       width = 14,
       height = 5,
       units = "in")
```

# Lowland tapir calf -dam over time

```{r}
# Select the data for the ex-situ mountain tapirs
meta_long <- metadata %>% 
  filter(Tapir == "Lowland") %>% 
  mutate(Date_collected = as.Date(.$Date_collected, format = "%m/%d/%Y")) %>% 
  mutate(months_since_start = time_length(interval(min(Date_collected), Date_collected), "months") + 1)

# Plot function
plot_Low <- function (data, variable) {
  ggplot(data = data, 
             aes(x = months_since_start, 
                 y = .data[[variable]], 
                 color = Animal_Code)) +
  geom_point(size = 4, shape = 17) +
  geom_line(aes(linetype = Group), 
            linewidth = 1.25) + 
  scale_linetype_manual(values = c(2,1), 
                        labels = c("Calf", "Adult"), 
                        name = "Age Group") + 
  scale_color_manual(name = "Animal", 
                     values = c("#c77cff","#00c08b")) +
  guides(color = guide_legend(
    override.aes = list(
      shape = c(17,17),
      linetype = c(2,1)
    ), 
    order = 2), 
    linetype = guide_legend(order = 1)) + 
  theme(title = element_text(size = 18),
        plot.title = element_text(hjust = 0.5),
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15),
        legend.text = element_text(size = 13), 
        legend.key.width = unit(2, "cm"))
}


# Richness
# Bacteria 
p1 <- plot_Low(data = meta_long, variable = "rich_bac") + 
  ylab("Species Richness") + 
  xlab("") + 
  ggtitle("Bacteria") 
p1
# Archaea 
p2 <- plot_Low(data = meta_long, variable = "rich_arch") + 
  ylab("") + 
  xlab("") + 
  ggtitle("Archaea") 
p2
# Fungi 
p3 <- plot_Low(data = meta_long, variable = "rich_fun") + 
  ylab("") + 
  xlab("") + 
  ggtitle("Fungi") 
p3

# Shannon
# Bacteria 
p4 <- plot_Low(data = meta_long, variable = "shannon_bac") + 
  ylab("Shannon Diversity") + 
  xlab("") + 
  ggtitle("") + 
  ylim(0,8) 
p4
# Archaea 
p5 <- plot_Low(data = meta_long, variable = "shannon_arch") + 
  ylab("") + 
  xlab("") + 
  ggtitle("") + 
  ylim(0,8)
p5
# Fungi
p6 <- plot_Low(data = meta_long, variable = "shannon_fun") + 
  ylab("") + 
  xlab("") + 
  ggtitle("") + 
  ylim(0,8)
p6

# Simpson
# Bacteria 
p7 <- plot_Low(data = meta_long, variable = "simpson_bac") + 
  ylab("Simpson Diversity") + 
  xlab("") + 
  ggtitle("") + 
  ylim(0,1)
p7
# Archaea
p8 <- plot_Low(data = meta_long, variable = "simpson_arch") + 
  ylab("") + 
  xlab("Calf age (month of life)") + 
  ggtitle("") + 
  ylim(0,1)
p8
# Fungi
p9 <- plot_Low(data = meta_long, variable = "simpson_fun") + 
  ylab("") + 
  xlab("") + 
  ggtitle("") +
  ylim(0,1)
p9


# Mega panel
((p1 | p2 | p3 ) / (p4 | p5 | p6) / (p7 | p8 | p9)) + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = "A", title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        legend.position = "right",
        plot.title= element_markdown())  


# Save the plots

# ggsave(file = "../Images/Manuscript (Frontiers)/Lowland_alpha_9.tiff", 
#        width = 1725,
#        height = 894,
#        units = "px",
#        dpi = 96.14)
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/Lowland_alpha_9.png", 
#        width = 1725,
#        height = 894,
#        units = "px",
#        dpi = 96.14)


# The panel we are actually going to use
((p4 + ggtitle("Bacteria")) | (p5 + ggtitle("Archaea") + xlab("Calf age (month of life)")) | (p6 + ggtitle("Fungi")) ) + 
  plot_layout(guides = 'collect') +
  plot_annotation(tag_levels = "A", title = "") & 
  theme(plot.tag = element_text(size = 25, face = "bold"),
        plot.tag.position = c(0.02, 1.02), 
        legend.position = "right",
        plot.title= element_markdown()) 

# Save the figure
# ggsave(file = "../Images/Manuscript (Frontiers)/Lowland_alpha_3_shannon.tiff", 
#        width = 14,
#        height = 5,
#        units = "in")
# 
# ggsave(file = "../Images/Manuscript (Frontiers)/Lowland_alpha_3_shannon.png", 
#        width = 14,
#        height = 5,
#        units = "in")
```


